---
title: "Homework 3"
author: "Maya Szymborski"
format: html
editor: visual
---

## Task 1: Conceptual Questions

**1) If your working directory is myfolder/homework/, what *relative* path would you specify to get the file located at myfolder/MyData.csv?**

../MyData.csv/

**2) What are the major benefits of using R projects?**

R projects allow you to use a single directory and workspace to easily collaborate with others.

**3) What is Git and what is GitHub?**

Git is a local version control software that is used to track changes within a code. GitHub is an online hosting service for Git-based projects that allows you to collaborate with others.

**4) What are the two main differences between a tibble and a data.frame?**

A `tibble` is an optimized version of a data frame that prints cleaner data. It is the main object of the `tidyverse` packaging. When printed, it only shows the first 10 rows of the data set and the data type of each column. When extracting a single column, it will return another `tibble`.

A `data.frame` is the standard `Base R` object. When printed, it shows the entire data set and does not show the data type. When extracting a single column, it will return a `vector`.

## Task 2 Reading Delimited Data

1)  Read the `glass.data` into R directly from the URL using functions from the tidyverse. Add column names. Print out the `tibble` (just call the object name).

```{r}
# install tidyverse package
# import data
library(tidyverse)
glass_data <- read_csv("https://www4.stat.ncsu.edu/online/datasets/glass.data", 
                       col_names = c("Id", "RI", "Na", "Mg", "Al", "Si", "K", "Ca", 
                                     "Ba", "Fe", "Type_of_Glass"))

# print tibble
glass_data
```

2)  Read the `yeast.data` into R directly from the URL using functions from the tidyverse. Add column names. Print out the `tibble` (just call the object name).

```{r}
# import data
yeast_data <- read_table("https://www4.stat.ncsu.edu/~online/datasets/yeast.data", 
                         col_names = c("seq_name", "mcg", "gvh", "alm", "mit", 
                                       "erl", "pox", "vac", "nuc", "class"))

# print tibble
yeast_data
```

## Task 3: Combining Excel and Delimited Data

3)  Import the `white-wine` data from the first sheet using the `readxl` package and print out the `tibble` (just call the object name).

```{r}
# import data
library(readxl)
white_wine_data <- read_excel("white-wine.xlsx", sheet = 1)

# print tibble
white_wine_data
```

4)  Read in the data from the 2nd sheet. This should return a data frame with one column containing alternative versions of the variable names.

```{r}
# import data
white_wine_data_2 <- read_excel("white-wine.xlsx", sheet = 2)

# print tibble
white_wine_data_2
```

-   Grab that column and overwrite the current column names (`colnames())` of your white wine `tibble`. Note that `colnames(object)` allows for ‘left hand side assignment’. That is, you can call it on the left and store an object in it to overwrite it. Similar to names(object) \<- c(...)

```{r}
# take the column from white_wine_data_2 and use as column names for white_wine_data

# check column names of wine_data first
colnames(white_wine_data)

# overwrite those column names
colnames(white_wine_data) <- white_wine_data_2[[1]]

# check new names
colnames(white_wine_data)
```

5)  Lastly, add a column to this data set to indicate the wines are white. That is, add a column that has values of ‘white’ for every observation. You can do this via the `$` operator or, if you know it already, the `mutate()` function from `dplyr`.

```{r}
# add 'type' column using `$` operator
white_wine_data$type <- "white"

# check tibble
white_wine_data
```

6)  There is a semi-colon delimited version of the red wine data set available at https://www4.stat.ncsu.edu/\~online/datasets/red-wine.csv. Read this in using the `readr` package. Be careful that the columns are read in as the correct type!

```{r}
library(readr) 

# import data 
red_wine_data <- read_delim("red-wine.csv", delim = ";") 

# print tibble 
red_wine_data
```

-   You should replace the variable names as done above

```{r}
# overwrite column names 
colnames(red_wine_data) <- white_wine_data_2[[1]] 

# check new names 
colnames(red_wine_data)
```

-   You should append a column denoting the type as “red”

```{r}
# add 'type' column 
red_wine_data$type <- "red" 

# check tibble 
red_wine_data
```

7)  Combine these two data sets into one data set. They both have the exact same columns so this is an easy append task!

-   Use the `dplyr::bind_rows()` function (see the help) to create one `tibble` containing all of the wine data.

```{r}
library(dplyr)

# combine both wine data sets 
wine_data <- bind_rows(white_wine_data, red_wine_data) 

# check & confirm both data sets combined
wine_data
table(wine_data$type)
```

## Task 4: Writing a Function

8)  Write a basic function (call it `getIQR`) that takes in a vector of numbers and returns the interquartile range (75th percentile - 25th percentile). You can use the `quantile` function. Remove the name on the returned vector (i.e. make sure the returned item doesn’t have a name like 75% or something like that). Allow your function to pass unnamed arguments to the `quantile` function.

```{r}
# the IQR is Q3 - Q1

getIQR <- function(x, ...) {
  q <- quantile(x, probs = c(0.25, 0.75), ...)
  IQR <- q[2] - q[1]
  names(IQR) <- NULL
  return(IQR)
}

# 'x' is the vector of numbers
# '...' allows unnamed arguments to pass through quantile()
# names(IQR) <- NULL removes names 
```

9)  Take the object from question 7 and apply the `getIRQ()` function to the `fixed_acidity` and `citric_acid` columns (two function calls).

```{r}
getIQR(wine_data$fixed_acidity)
getIQR(wine_data$citric_acid)
```

10) Write another function (call it `getCoefVar`) that returns the sample coefficient of variation, defined as the sample standard deviation divided by the absolute value of the sample mean. Allow the function to pass unnamed arguments to sd and mean.

```{r}
getCoefVar <- function(x, ...) {
  sd <- sd(x, ...)
  mean <- mean(x, ...)
  cv <- sd/abs(mean)
  return(cv)
}
```

11) Repeat question 9 using the `getCoefVar()` function.

```{r}
getCoefVar(wine_data$fixed_acidity)
getCoefVar(wine_data$citric_acid)
```

12) Create a function (call it `getSummaries`) that takes in a vector and returns both the IQR and the coefficient of variation. Do not rewrite your above two functions, call them. When returning your values, give them appropriate names. (This would be called a `wrapper` function.)

```{r}
# use getIQR() and getCoefVar() inside this function

getSummaries <- function(x, ...) {
  IQR <- getIQR(x, ...)
  CV <- getCoefVar(x, ...)
  return(c(IQR = IQR, CoefVar = CV)) # gives appropriate names
}
```

13) Repeat question 2 with the `getSummaries()` function.

```{r}
getSummaries(wine_data$fixed_acidity)
getSummaries(wine_data$citric_acid)
```

14) Write your own `infix` function that can be used in replace of the `cbind` function (feel free to call the `cbind` function in your function :) Check that the function works on the two data frames below (`df1` & `df2`) and then all three objects (`df1`, `df2`, `vec1`)!

```{r}
# an infix funciton goes between arguments 
# and uses symbols on both sides (%symbol%)

# create infix function, using `` to define it
`%cbind%` <- function(x, y) {
  cbind(x, y)
}

# %cbind% operator takes object on left as 'x' and object on right as 'y'

df1 <- data.frame(unif = runif(10), norm = rnorm(10))

df2 <- data.frame(gamma = rgamma(10, shape = 1), 
                  beta = rbeta(10, shape1 = 5, shape2 = 1))

vec1 <- rnorm(10, mean = 10)

# check on data frames and vector

df1 %cbind% df2 # same as saying cbind(x, y)

df1 %cbind% df2 %cbind% vec1
```
