---
title: "Homework 3"
author: "Maya Szymborski"
format: html
editor: visual
---

## Task 1: Conceptual Questions

**1) If your working directory is myfolder/homework/, what *relative* path would you specify to get the file located at myfolder/MyData.csv?**

../MyData.csv/

**2) What are the major benefits of using R projects?**

R projects allow you to use a single directory and workspace to easily collaborate with others.

**3) What is Git and what is GitHub?**

Git is a local version control software that is used to track changes within a code. GitHub is an online hosting service for Git-based projects that allows you to collaborate with others.

**4) What are the two main differences between a tibble and a data.frame?**

A `tibble` is an optimized version of a data frame that prints cleaner data. It is the main object of the `tidyverse` packaging. When printed, it only shows the first 10 rows of the data set and the data type of each column. When extracting a single column, it will return another `tibble`.

A `data.frame` is the standard `Base R` object. When printed, it shows the entire data set and does not show the data type. When extracting a single column, it will return a `vector`.

## Task 2 Reading Delimited Data

1)  Read the `glass.data` into R directly from the URL using functions from the tidyverse. Add column names. Print out the `tibble` (just call the object name).

```{r}
# install tidyverse package
# import data
library(tidyverse)
glass_data <- read_csv("https://www4.stat.ncsu.edu/online/datasets/glass.data", 
                       col_names = c("Id", "RI", "Na", "Mg", "Al", "Si", "K", "Ca", 
                                     "Ba", "Fe", "Type_of_Glass"))

# print tibble
glass_data
```

2)  Read the `yeast.data` into R directly from the URL using functions from the tidyverse. Add column names. Print out the `tibble` (just call the object name).

```{r}
# import data
yeast_data <- read_table("https://www4.stat.ncsu.edu/~online/datasets/yeast.data", 
                         col_names = c("seq_name", "mcg", "gvh", "alm", "mit", 
                                       "erl", "pox", "vac", "nuc", "class"))

# print tibble
yeast_data
```

## Task 3: Combining Excel and Delimited Data

3)  Import the `white-wine` data from the first sheet using the `readxl` package and print out the `tibble` (just call the object name).

```{r}
# import data
library(readxl)
wine_data <- read_excel("white-wine.xlsx", sheet = 1)

# print tibble
wine_data
```

4)  Read in the data from the 2nd sheet. This should return a data frame with one column containing alternative versions of the variable names.

```{r}
# import data
wine_data_2 <- read_excel("white-wine.xlsx", sheet = 2)

# print tibble
wine_data_2
```

- Grab that column and overwrite the current column names (`colnames())` of your white wine `tibble`. Note that `colnames(object)` allows for ‘left hand side assignment’. That is, you can call it on the left and store an object in it to overwrite it. Similar to names(object) \<- c(...)

```{r}
# take the column from wine_data_2 and use as column names for wine_data

# check column names of wine_data first
colnames(wine_data)

# overwrite those column names
colnames(wine_data) <- wine_data_2[[1]]

# check new names
colnames(wine_data)
```

5)  Lastly, add a column to this data set to indicate the wines are white. That is, add a column that has values of ‘white’ for every observation. You can do this via the `$` operator or, if you know it already, the `mutate()` function from `dplyr`.

```{r}
# add 'type' column using `$` operator
wine_data$type <- "white"

# check tibble
wine_data
```

6)  There is a semi-colon delimited version of the red wine data set available at https://www4.stat.ncsu.edu/\~online/datasets/red-wine.csv. Read this in using the readr package. Be careful that the columns are read in as the correct type!

```{r}

```

- You should replace the variable names as done above

```{r}

```

- You should append a column denoting the type as “red”

```{r}

```

7)  Combine these two data sets into one data set. They both have the exact same columns so this is an easy append task!

```{r}

```

- Use the dplyr::bind_rows() function (see the help) to create one tibble containing all of the wine data.

```{r}

```

## Task 4: Writing a Function

8)  Write a basic function (call it getIQR) that takes in a vector of numbers and returns the interquartile range (75th percentile - 25th percentile). You can use the quantile function. Remove the name on the returned vector (i.e. make sure the returned item doesn’t have a name like 75% or something like that). Allow your function to pass unnamed arguments to the quantile function.

```{r}

```

9)  Take the object from question 7 and apply the getIRQ() function to the fixed_acidity and citric_acid columns (two function calls).

```{r}

```

10) Write another function (call it getCoefVar) that returns the sample coefficient of variation, defined as the sample standard deviation divided by the absolute value of the sample mean. Allow the function to pass unnamed arguments to sd and mean.

```{r}

```

11) Repeat question 9 using the getCoefVar() function.

```{r}

```

12) Create a function (call it getSummaries) that takes in a vector and returns both the IQR and the coefficient of variation. Do not rewrite your above two functions, call them. When returning your values, give them appropriate names. (This would be called a wrapper function.)

```{r}

```

13) Repeat question 2 with the getSummaries() function.

```{r}

```

14) Write your own infix function that can be used in replace of the cbind function (feel free to call the cbind function in your function :) Check that the function works on the two data frames below (df1 & df2) and then all three objects (df1, df2, vec1)!

```{r}

```
