---
title: "HW3_B"
author: "Daria Smyslova"
format: html
editor: visual
---

### Task 1: Conceptual Questions

#### 1) If your working directory is myfolder/homework/, what relative path would you specify to get the file located at myfolder/MyData.csv?

Since the working directory is `myfolder/homework/`, `..` moves up one directory to `myfolder/`, where `MyData.csv` is located.

#### 2) What are the major benefits of using R projects?

R Projects keep all files for a project together, automatically establish the project directory, make relative file paths easier to use, and help make analyses organized, portable, and reproducible.

#### 3) What is Git and what is GitHub?

Git is a version control system that tracks changes to files and allows you to return to earlier versions or collaborate on code. GitHub is an online platform that hosts Git repositories, making it easier to store, share, and collaborate on projects.

#### 4) What are the two main differences between a tibble and a data.frame?

A tibble is a newer version of a data frame in R. One difference is how they print: tibbles only show the rows and columns that fit on the screen, while data frames may print the whole dataset. Another difference is when you select a single column. A tibble usually keeps it as a tibble, while a data frame may turn it into a vector.

### Task 2 Reading Delimited Data

#### 1) Read this data into R directly from the URL using functions from the tidyverse. Notice that the data doesn’t include column names - add those (in a manner of your choosing). Print out the tibble (just call the object name).

```{r}
#glass data
library(tidyverse)

glass <- read_csv(
  "https://www4.stat.ncsu.edu/~online/datasets/glass.data",
  col_names = c(
    "Id", "RI", "Na", "Mg", "Al", "Si",
    "K", "Ca", "Ba", "Fe", "Type"
  )
)

glass
```

#### 2) Read this data into R directly from the URL using functions from the tidyverse. Notice that the data doesn’t include column names - add those (in a manner of your choosing). Print out the tibble (just call the object name).

```{r}
#yeast data
yeast <- read_table(
  "https://www4.stat.ncsu.edu/~online/datasets/yeast.data",
  col_names = c(
    "seq_name", "mcg", "gvh", "alm", "mit",
    "erl", "pox", "vac", "nuc", "class"
  )
)

yeast
```

### Task 3: Combining Excel and Delimited Data

#### 3) There is an excel version of the white wine data set available at https://www4.stat.ncsu.edu/\~online/datasets/white-wine.xlsx.

##### • Import the data from the first sheet using the readxl package

```{r}
library(readxl)

white_wine <- read_excel("white-wine.xlsx", sheet = 1)
```

##### • Print out the tibble (just call the object name)

```{r}
white_wine
```

#### 4) Read in the data from the 2nd sheet. This should return a data frame with one column containing alternative versions of the variable names.

```{r}
wine_names <- read_excel("Homework 3/white-wine.xlsx", sheet = 2)
wine_names
```

##### • Grab that column and overwrite the current column names (colnames()) of your white wine tibble. Note that colnames(object) allows for ‘left hand side assignment’. That is, you can call it on the left and store an object in it to overwrite it. Similar to names(object) \<- c(...)

```{r}
colnames(white_wine) <- wine_names[[1]]

white_wine
```

#### 5) Lastly, add a column to this data set to indicate the wines are white. That is, add a column that has values of ‘white’ for every observation. You can do this via the \$ operator or, if you know it already, the mutate() function from dplyr.

```{r}
white_wine$type <- "white"

white_wine
```

#### 6) There is a semi-colon delimited version of the red wine data set available at https://www4.stat.ncsu.edu/\~online/datasets/red-wine.csv. Read this in using the readr package. Be careful that the columns are read in as the correct type!

```{r}
library(readr)

red_wine <- read_delim(
  "Homework 3/red-wine.csv",
  delim = ";"
)

red_wine
```

##### • You should replace the variable names as done above

```{r}
colnames(red_wine) <- wine_names[[1]]

red_wine
```

##### • You should append a column denoting the type as “red”

```{r}
red_wine$type <- "red"

red_wine
```

#### 7) Combine these two data sets into one data set. They both have the exact same columns so this is an easy append task!

```{r}
wine <- dplyr::bind_rows(white_wine, red_wine)
```

##### •Use the dplyr::bind_rows() function (see the help) to create one tibble containing all of the wine data.

```{r}
wine
```

### Task 4: Writing a Function

#### 8) Write a basic function (call it getIQR) that takes in a vector of numbers and returns the interquartile range (75th percentile - 25th percentile). You can use the quantile function. Remove the name on the returned vector (i.e. make sure the returned item doesn’t have a name like 75% or something like that). Allow your function to pass unnamed arguments to the quantile function.

```{r}
getIQR <- function(x, ...) {
  q75 <- quantile(x, 0.75, ...)
  q25 <- quantile(x, 0.25, ...)
  
  unname(q75 - q25)
}
```

#### 9) Take the object from question 7 and apply the getIRQ() function to the fixed_acidity and citric_acid columns (two function calls).

```{r}
getIQR(wine$fixed_acidity)

getIQR(wine$citric_acid)
```

#### 10) Write another function (call it getCoefVar) that returns the sample coefficient of variation, defined as the sample standard deviation divided by the absolute value of the sample mean. Allow the function to pass unnamed arguments to sd and mean.

```{r}
getCoefVar <- function(x, ...) {
  sd(x, ...) / abs(mean(x, ...))
}
```

#### 11) Repeat question 9 using the getCoefVar() function.

```{r}
getCoefVar(wine$fixed_acidity)

getCoefVar(wine$citric_acid)
```

#### 12) Create a function (call it getSummaries) that takes in a vector and returns both the IQR and the coefficient of variation. Do not rewrite your above two functions, call them. When returning your values, give them appropriate names. (This would be called a wrapper function.)

```{r}
getSummaries <- function(x, ...) {
  c(
    IQR = getIQR(x, ...),
    CoefVar = getCoefVar(x, ...)
  )
}
```

#### 13) Repeat question 2 with the getSummaries() function.

```{r}
getSummaries(wine$fixed_acidity)

getSummaries(wine$citric_acid)
```

#### 14) Write your own infix function that can be used in replace of the cbind function (feel free to call the cbind function in your function :) Check that the function works on the two data frames below (df1 & df2) and then all three objects (df1, df2, vec1)!

```{r}
# Create two data frames and one vector for testing the infix function
df1 <- data.frame(
  unif = runif(10),
  norm = rnorm(10)
)

df2 <- data.frame(
  gamma = rgamma(10, shape = 1),
  beta = rbeta(10, shape1 = 5, shape2 = 1)
)

vec1 <- rnorm(10, mean = 10)
```

```{r}
#run infix function
`%c%` <- function(x, y) {
  cbind(x, y)
}
```

```{r}
# Test the infix function with two data frames
df1 %c% df2

# Test the infix function with two data frames and a vector
df1 %c% df2 %c% vec1
```
