---
title: "Homework #2"
author: "Michael Costello"
format: html
editor: visual
---

#Task 1: Basic Vector Practice

1.  Create two vectors of the pre- and post-measurements:

```{r}
pre_bp <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)

post_bp <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)
```

2.  Assign names to the vector elements:

```{r}
names(pre_bp)  <- paste("Subject", 1:20, sep = "_")
names(post_bp) <- paste("Subject", 1:20, sep = "_")
```

3.  Calculate the change (decrease?) in blood pressure accross all patients:

```{r}
diff_bp <- pre_bp - post_bp
diff_bp
```

4.  Find the average decrease in BP across all patients:

```{r}
mean(diff_bp)
```

5.  Patients who experienced a deceased in BP after treatment:

```{r}
improved_bp <- which(diff_bp > 0, useNames = TRUE)
```

6.  Create a new vector object with only improvers and their BPs:

```{r}
improvers <- diff_bp[improved_bp]
```

7.  Find the mean of the improvers BPs:

```{r}
mean(improvers)
```

#Task 2: Basic Data Frame Practice

1.  Create a frame object with four columns using the 4 vectors from Task 1:

```{r}
patient_name <- names(pre_bp)
bp_df <- data.frame(patient_name, pre_bp, post_bp, diff_bp )
#Meh, I don't love the double naming here. Thoughts?
bp_df
```

2.  Return only rows with negative BP:

```{r}
subset(bp_df, diff_bp < 0)

```

3.  Add new col to df with TRUE if the post_bp is below 120.

```{r}
bp_df$low_bp <- ifelse(bp_df$post_bp < 120, TRUE, "") #Should we list BPs >= 119 as FALSE?
bp_df
```

4.  Print the whole data frame.

```{r}
knitr::kable(bp_df)
```

#Task 3: List practice

1.  Create a new df similar to the one from Task 1.

```{r}
pre_bp_control <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)
post_bp_control   <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)
diff_bp_control <- pre_bp_control - post_bp_control
bp_df_control <- data.frame(pre_bp_control, post_bp_control, diff_bp_control)
bp_df_control
```

2.  Create a list with two elements (treatment (the first df), and placebo (the second df))

```{r}
my_list <- list(treatment = bp_df, placebo = bp_df_control)
my_list
```

3.  Access the first list 3 different ways.

```{r}
my_list[1]
my_list[[1]]
my_list$treatment
```

4.  Call the Placebo's pre-treatment BP in a single function.

```{r}
my_list$placebo$pre_bp_control
```

#Task 4: Control Flow Practice

1.  Create a variable called "status" in each of the two data frames.

```{r}
my_list$treatment$status <- character(20)
my_list$placebo$status   <- character(10)

```

2.  Create categories for BP in the status variable, for the treatment data frame.

```{r}
		for(i in 1:20){
  		  if (my_list$treatment$post_bp[i] > 130) {
  		  my_list$treatment$status[i] = "high"
  		  #starting at 130+ and then working my way down because higher numbers won't see the 2nd and 3rd else statements.
  		} else if (my_list$treatment$post_bp[i] > 120) {
  		  my_list$treatment$status[i] = "borderline"
  		} else {
  		  my_list$treatment$status[i] = "optimal"
  		}
		}
my_list$treatment
```

3.  Repeat for the placebo data frame.

```{r}
		for(i in 1:10){
  		  if (my_list$placebo$post_bp[i] > 130) {
  		  my_list$placebo$status[i] = "high"
  		  #starting at 130+ and then working my way down because higher numbers won't see the 2nd and 3rd else statements.
  		} else if (my_list$placebo$post_bp[i] > 120) {
  		  my_list$placebo$status[i] = "borderline"
  		} else {
  		  my_list$placebo$status[i] = "optimal"
  		}
		}
my_list$placebo
```
