---
title: "Homework 2"
author: "Maya Szymborski"
format: html
editor: visual
---

## Task 1: Basic Vector Practice

1. Create two vectors of pre- and post-measurements from a medical experiment on blood pressure 

```{r}
# create two vectors of pre-measurements and post-measurements using c() function

pre_bp <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)

post_bp <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)
```

2. Assign `names` to the vector elements using `paste()` function

```{r}
# give a name to each vector element using paste() function

names(pre_bp) <- paste("Subject", 1:20, sep = "_")

names(post_bp) <- paste("Subject", 1:20, sep = "_")

pre_bp
post_bp
```

3. Calculate the change in blood pressure for each patient

```{r}
# calculate change by subtracting post-measurements from pre-measurements
# a positive value is a decrease in blood pressure 
# save calculation as a new object

diff_bp <- pre_bp - post_bp
diff_bp
```

4. Calculate the average decrease in blood pressure across all patients

```{r}
# calculate average decrease in blood pressure 
mean(diff_bp)
```

5. Determine which patients experienced a decrease in blood pressure after treatment. 

```{r}
# use which() function to determine which subjects had a 
# decrease in blood pressure after treatment

which(diff_bp > 0)
```

6. Subset the vector of differences to only return those that have a positive change. 

```{r}
# subset diff_meas vector to only return vector of elements 
# with a positive change (decrease in blood pressure)
# create object containing positive changes 
bp_decrease <- diff_bp[diff_bp > 0]
bp_decrease
```

7. Calculate the average decrease in blood pressure for subjects where blood pressure decreased. 

```{r}
# calculate average decrease in blood pressure 
mean(bp_decrease)
```


## Task 2: Basic Data Frame Practice

1. Create a data frame object with four columns: patients, pre_bp, post_bp, and diff_bp

```{r}
# create vector for patients
patients <- names(pre_bp)

# create data frame with four columns using data.frame() function 
bp_df <- data.frame(patients, pre_bp, post_bp, diff_bp)
bp_df
```
 
2. Return only rows where the `diff_bp` column is negative. 

```{r}
# only show rows with a negative blood pressure difference using subset() function 
# means blood pressure increased 
subset(bp_df, diff_bp < 0)

# or can use brackets []
bp_df[bp_df$diff_bp < 0, ]
```

3. Add a new column to the data frame corresponding to `TRUE` if the `post_bp` is less than 120.

```{r}
# add a new column where post_bp is less than 120
bp_df$post_less_120 <- bp_df$post_bp < 120

# check that the column is added and shows correctly
bp_df
```

4. Print the data frame out nicely 

```{r}
knitr::kable(bp_df)
```

