---
title: "HW2_Group3"
format: html
editor: visual
---
**Group solution note**
**Task 1: Basic Vector Practice:** Members A, B, C and D used the same general approach for creating and naming the vectors and calculating the blood pressure differences. Some solutions differed in whether intermediate objects were created before subsetting. For the final version, we retained the pos object created with which(diff_bp \> 0) because it makes the connection between identifying positive differences in Question 5 and using those positions for subsetting in Questions 6 and 7 explicit.

**Task 2: Basic Data Frame Practice:** Members A, B, C and D used slightly different approaches to constructing and manipulating the data frame. One difference was whether logical conditions were first stored as separate vectors or assigned directly to a data frame column. For the final version, we used `bp_df1$less_120 <- bp_df1$post_bp < 120` because the comparison already produces a logical vector of the appropriate length, so creating a separate intermediate object is unnecessary. We also used bracket subsetting for the negative differences to practice accessing a column directly from the data frame, as requested in the assignment.

**Task 3: List Practice:** Members B and D had some differences in how intermediate objects were created when constructing the placebo data frame. For the final version, we retained the more explicit construction with separate vectors before combining them into the data frame because it parallels the workflow used in Tasks 1 and 2 and makes each step easier to follow. For accessing elements of the list, both members used multiple valid forms of Base R syntax. We kept `[ ]`, `[[ ]]`, and `$` in the final solution because the task specifically asks for three different methods and they demonstrate an important distinction: `[ ]` returns a sub-list, whereas `[[ ]]` and `$` extract the element itself. For accessing the placebo `pre_bp` column, several valid approaches were identified. We retained `bp_exp$placebo$pre_bp` as the clearest final answer because it directly extracts the requested column as a vector in one line.

**Task 4: Control Flow Practice:** Members B and D followed the same general control-flow logic, with minor differences in how the loops and conditions were written. For the final version, we used an `if` / `else if` / `else` structure that checks the thresholds in ascending order. Once values at or below 120 are classified as `"optimal"`, the next condition only needs to check whether the value is at or below 130; any remaining value must therefore be `"high"`. This keeps the conditions concise while matching the blood pressure categories specified in the assignment. We also retained separate loops for the treatment and placebo data frames rather than combining the process because the assignment specifically asks for the procedure to be repeated for the placebo data.

### Task 1: Basic Vector Practice

#### 1) Create two vectors using the `c()` function

```{r}
# create vector with pre measurements using combine function
pre_bp <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)

# create vector with post measurements using combine function
post_bp <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)
```

#### 2) Assign names to the vector elements using the `paste()` function

```{r}
# create names for the 20 subjects using a character vector
names(pre_bp) <- paste("Subject", 1:20, sep = "_")
names(post_bp) <- paste("Subject", 1:20, sep = "_")

pre_bp
post_bp
```

#### 3) Calculate change in blood pressure

```{r}
#subtract the post treatment level from pre treatment level for each subject and store as vector
diff_bp <- (pre_bp - post_bp)
diff_bp
```

This shows that two numeric vectors of equal length can be subtracted directly, and R performs the operation element by element. Of the 20 subjects, all but three showed decreases in blood pressure.

#### 4) Calculate average change in blood pressure

```{r}
mean(diff_bp)
```

The average decrease in blood pressure was 17.

#### 5) Identify which subjects showed a decrease in blood pressure

```{r}
# we use a logical operator to determine positive values
which(diff_bp > 0)
```

The output gives the indices and names associated with positive changes.

#### 6) Subset the vector of differences for only positive differences

```{r}
#one way of doing this is to create a position vector of positive differences
pos <- which(diff_bp > 0)

# returns the elements of the positions listed in pos
diff_bp[pos]
```

We use `[]` in R to subset or select certain elements from an object.

#### 7) Calculate average change in blood pressure for positive differences

```{r}
mean(diff_bp[pos])
```

The average decrease in blood pressure was 20.65.

### Task 2: Basic Data Frame Practice

#### 1) Create a data frame with four columns

```{r}
# create a separate character vector for the subject names
patient <- names(pre_bp)

#create data frame using `data.frame()` function
bp_df1 <- data.frame(patient, pre_bp, post_bp, diff_bp)
bp_df1
```

The big teaching here is that the subject names, up until this step, were stored as the `names` attribute of the vectors and not as a separate column. In order to use them as the `patient` column in the data frame we first extracted them into their own character vector. We also note that vector names can automatically become data frame row names, but that row names are not the same thing as a data frame column.

#### 2) Return data frame rows where `diff` column is negative

```{r}
#from the data frame, keep rows where `diff` is negative
bp_df1[bp_df1$diff_bp < 0, ]
```

This confirms that of the 20 subjects, three of them saw negative differences between the pre and post treatments.

#### 3) Add a new column using a condition

```{r}
# identify in data frame where post is less than 120 
bp_df1$post < 120 # this creates a logical vector

#create a new column with the logical vector
bp_df1$less_120 <- bp_df1$post < 120
```

The key teaching here is that the comparison `bp_df1$post_bp < 120`produced a logical vector, but it did not need to be stored as a separate vector object in order to assign it directly to the data frame column. Alternatively, we could have saved the logical vector as a separate vector object and then assigned it to the data frame, but this would have introduced an extra step.

#### 4) Print the data frame out nicely

```{r}
knitr::kable(bp_df1)
```

### Task 3: List Practice

#### 1) Create a new dataframe

```{r}
# create two new vectors of pre-measurements and post-measurements of the second experiment 
pre_bp2 <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)
post_bp2 <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)
# create pre_bp2 and post_bp2 objects
names(pre_bp2) <- paste("Subject", 1:10, sep = "_")
names(post_bp2) <- paste("Subject", 1:10, sep = "_")

pre_bp2
post_bp2

# create diff_bp2 object
diff_bp2 <- pre_bp2 - post_bp2
diff_bp2

# create post_less_120_2 object
post_less_120_2 <- post_bp2 < 120
post_less_120_2

# create patients2 object
patients2 <- names(pre_bp2)
patients2

# create second data frame that is similar to the previous
bp_df2 <- data.frame(
  patients = patients2,
  pre_bp = pre_bp2,
  post_bp = post_bp2,
  diff_bp = diff_bp2,
  post_under_120 = post_less_120_2
)
bp_df2

# print data frame nicely
knitr::kable(bp_df2)
```

#### 2) Create and store a list with two elements

-   1st element named **treatment** and contains the first data frame
-   2nd element named **placebo** and contains the second data frame

```{r}
# create treatment and placebo objects for each corresponding dataframe
treatment <- bp_df1
placebo <- bp_df2

# create list with treatment and placebo using list()
bp_exp <- list(treatment = treatment, placebo = placebo)
bp_exp
```

#### 3) Access the first list element using three different types of syntax.

```{r}
# first access using brackets [], returns a list containing the named element
bp_exp[1]

# second access using double brackets [[]], returns the actual data frame 
bp_exp[[1]]

# third access using $, also returns just the data frame itself
bp_exp$treatment
```

#### 4) In one line, access the `placebo` data frame, `pre_bp2` column.

```{r}
# use [] and [[]] to access the pre_bp2 column as a data frame
bp_exp[[2]] [2:2]

# or use $ and [] to do the same 
bp_exp$placebo[2]

# or use just $ to return the vector only
bp_exp$placebo$pre_bp
```

### Task 4: Control Flow Practice

Characterize the post-treatment/placebo blood pressure as **optimal (\<= 120)**, **borderline (120 \< bp \<= 130)**, and **high (\> 130)**.

#### 1) Create a new column in each data frame called status.

```{r}
# create a new "status" column in each data frame
bp_exp$treatment$status <- character(20)
bp_exp$placebo$status <- character(10)

# check that the new columns are added
bp_exp
```

#### 2) For the non-placebo data frame (within the list), create a `for loop` and use `if/then/else` logic to create the status column's values.

```{r}
# use a for loop and if/then/else logic to create characterizations for the blood pressure measurements 

for(i in 1:20){
  if (bp_exp$treatment$post_bp[i] <= 120) {
    bp_exp$treatment$status[i] <- "optimal"
    
  } else if (bp_exp$treatment$post_bp[i] <= 130) {
    bp_exp$treatment$status[i] <- "borderline"
    
  } else {
    bp_exp$treatment$status[i] <- "high"
  }
}

# check
bp_exp$treatment
```

#### 3) Repeat for the placebo data frame (within the list).

```{r}
# repeat as before but with the placebo data frame

for(i in 1:10){
  if (bp_exp$placebo$post_bp[i] <= 120) {
    bp_exp$placebo$status[i] <- "optimal"
    
  } else if (bp_exp$placebo$post_bp[i] <= 130) {
    bp_exp$placebo$status[i] <- "borderline"
    
  } else {
    bp_exp$placebo$status[i] <- "high"
  }
}

# check
bp_exp$placebo

# check that they both show up on the list
bp_exp
```
