---
title: "Homework 2 - Member A"
author: "Alfonso Straffon"
format: html
editor: visual
---


## Task 1: Basic Vector Practice

### 1) Create two vectors using the `c()` function

```{r}
# create vector with pre measurements using combine function
pre_bp <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)

# create vector with post measurements using combine function
post_bp <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)

```

### 2) Assign names to the vector elements using the `paste()` function 

```{r}

# create names for the 20 subjects using a character vector
names(pre_bp) <- paste("Subject", 1:20, sep = "_")
names(post_bp) <- paste("Subject", 1:20, sep = "_")

pre_bp
post_bp

```
### 3) Calculate change in blood pressure

```{r}
#subtract the post treatment level from pre treatment level for each subject and store as vector
diff_bp <- (pre_bp - post_bp)
diff_bp

```
This shows that two numeric vectors of equal length can be subtracted directly, and 
R performs the operation element by element. Of the 20 subjects, all but three showed 
decreases in blood pressure. 

### 4) Calculate average change in blood pressure

```{r}

mean(diff_bp)
```
The average decrease in blood pressure was 17.

### 5) Identify which subjects showed a decrease in blood pressure

```{r}
# we use a logical operator to determine positive values
which(diff_bp > 0)

```
The output gives the indices and names associated with positive changes.

### 6) Subset the vector of differences for only positive differences

```{r}
#one way of doing this is to create a position vector of positive differences
pos_indice <- which(diff_bp > 0)

# returns the elements of the positions listed in pos_indice
diff_bp[pos_indice]

```
We use `[]` in R to subset or select certain elements from an object. 

### 7) Calculate average change in blood pressure for positive differences

```{r}

mean(diff_bp[pos_indice])
```
The average decrease in blood pressure was 20.65.

## Task 2: Basic Data Frame Practice

## Task 3: List Practice

## Task 4: Control Flow Practice