---
title: "Homework 2 - Member A"
author: "Alfonso Straffon"
format: html
editor: visual
---

## Task 1: Basic Vector Practice

### 1) Create two vectors using the `c()` function

```{r}
# create vector with pre measurements using combine function
pre_bp <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)

# create vector with post measurements using combine function
post_bp <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)

```

### 2) Assign names to the vector elements using the `paste()` function

```{r}

# create names for the 20 subjects using a character vector
names(pre_bp) <- paste("Subject", 1:20, sep = "_")
names(post_bp) <- paste("Subject", 1:20, sep = "_")

pre_bp
post_bp

```

### 3) Calculate change in blood pressure

```{r}
#subtract the post treatment level from pre treatment level for each subject and store as vector
diff_bp <- (pre_bp - post_bp)
diff_bp

```

This shows that two numeric vectors of equal length can be subtracted directly, and R performs the operation element by element. Of the 20 subjects, all but three showed decreases in blood pressure.

### 4) Calculate average change in blood pressure

```{r}

mean(diff_bp)
```

The average decrease in blood pressure was 17.

### 5) Identify which subjects showed a decrease in blood pressure

```{r}
# we use a logical operator to determine positive values
which(diff_bp > 0)

```

The output gives the indices and names associated with positive changes.

### 6) Subset the vector of differences for only positive differences

```{r}
#one way of doing this is to create a position vector of positive differences
pos_indice <- which(diff_bp > 0)

# returns the elements of the positions listed in pos_indice
diff_bp[pos_indice]

```

We use `[]` in R to subset or select certain elements from an object.

### 7) Calculate average change in blood pressure for positive differences

```{r}

mean(diff_bp[pos_indice])
```

The average decrease in blood pressure was 20.65.

## Task 2: Basic Data Frame Practice

### 1) Create a data frame with four columns

```{r}
# create a separate character vector for the subject names
patient <- names(pre_bp)

#create data frame using `data.frame()` function
bp_df <- data.frame(patient, pre_bp, post_bp, diff_bp)
bp_df
```
The big teaching here is that the subject names, up until this step, were stored as 
the `names` attribute of the vectors and not as a separate column. In order to use
them as the `patient` column in the data frame we first extracted them into their own 
character vector. We also note that vector names can automatically become data frame 
row names, but that row names are not the same thing as a data frame column.

### 2) Return data frame rows where `diff_bp` column is negative

```{r}

#from the data frame, keep rows where `bp_diff` is negative
bp_df[bp_df$diff_bp < 0, ]
```
This confirms that of the 20 subjects, three of them saw negative differences between
the pre and post treatments.

### 3) Add a new column using a condition

```{r}
# identify in data frame where post_bp is less than 120 
bp_df$post_bp < 120 # this creates a logical vector

#create a new column with the logical vector
bp_df$less_120 <- bp_df$post_bp < 120
```
The key teaching here is that the comparison `bp_df$post_bp < 120`produced a logical
vector, but it did not need to be stored as a separate vector object in order to assign
it directly to the data frame column. Alternatively, we could have saved the logical 
vector as a separate vector object and then assigned it to the data frame, but this 
would have introduced an extra step.

### 4) Print the data frame in document

```{r}
 knitr::kable(bp_df)
```


## Task 3: List Practice



## Task 4: Control Flow Practice
