---
title: "Homework 2"
author: "Maya Szymborski"
format: html
editor: visual
---

## Task 1: Basic Vector Practice

1. Create two vectors of pre- and post-measurements from a medical experiment on blood pressure 

```{r}
# create two vectors of pre-measurements and post-measurements using c() function

pre_bp1 <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)

post_bp1 <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)
```

2. Assign `names` to the vector elements using `paste()` function

```{r}
# give a name to each vector element using paste() function

names(pre_bp1) <- paste("Subject", 1:20, sep = "_")

names(post_bp1) <- paste("Subject", 1:20, sep = "_")

pre_bp1
post_bp1
```

3. Calculate the change in blood pressure for each patient

```{r}
# calculate change by subtracting post-measurements from pre-measurements
# a positive value is a decrease in blood pressure 
# save calculation as a new object

diff_bp1 <- pre_bp1 - post_bp1
diff_bp1
```

4. Calculate the average decrease in blood pressure across all patients

```{r}
# calculate average decrease in blood pressure 
mean(diff_bp1)
```

5. Determine which patients experienced a decrease in blood pressure after treatment. 

```{r}
# use which() function to determine which subjects had a 
# decrease in blood pressure after treatment

which(diff_bp1 > 0)
```

6. Subset the vector of differences to only return those that have a positive change. 

```{r}
# subset diff_meas vector to only return vector of elements 
# with a positive change (decrease in blood pressure)
# create object containing positive changes 
bp_decrease1 <- diff_bp1[diff_bp1 > 0]
bp_decrease1
```

7. Calculate the average decrease in blood pressure for subjects where blood pressure decreased. 

```{r}
# calculate average decrease in blood pressure 
mean(bp_decrease1)
```


## Task 2: Basic Data Frame Practice

1. Create a data frame object with four columns: **patients**, **pre_bp**, **post_bp**, and **diff_bp**

```{r}
# create vector for patients
patients1 <- names(pre_bp1)

# create data frame with four columns using data.frame() function 
bp_df1 <- data.frame(patients1, pre_bp1, post_bp1, diff_bp1)
bp_df1
```
 
2. Return only rows where the `diff_bp` column is negative. 

```{r}
# only show rows with a negative blood pressure difference using subset() function 
# means blood pressure increased 
subset(bp_df1, diff_bp1 < 0)

# or can use brackets []
bp_df1[bp_df1$diff_bp1 < 0, ]
```

3. Add a new column to the data frame corresponding to `TRUE` if the `post_bp` is less than 120.

```{r}
# add a new column where post_bp is less than 120
bp_df1$post_less_120_1 <- bp_df1$post_bp1 < 120

# check that the column is added and shows correctly
bp_df1
```

4. Print the data frame out nicely 

```{r}
knitr::kable(bp_df1)
```

## Task 3: List Practice

1. Create a new data frame with new data that is similar to the previous data frame. 

```{r}
# create a similar data frame to the first one (Task 2)
# repeating steps from Task 1 and 2

# create two new vectors of pre-measurements and post-measurements of the second experiment 
pre_bp2 <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)
post_bp2 <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)

# create pre_bp2 and post_bp2 objects
names(pre_bp2) <- paste("Subject", 1:10, sep = "_")
names(post_bp2) <- paste("Subject", 1:10, sep = "_")

pre_bp2
post_bp2

# create diff_bp2 object
diff_bp2 <- pre_bp2 - post_bp2
diff_bp2

# create post_less_120_2 object
post_less_120_2 <- post_bp2 < 120
post_less_120_2

# create patients2 object
patients2 <- names(pre_bp2)
patients2

# create second data frame that is similar to the previous
bp_df2 <- data.frame(
  patients = patients2,
  pre_bp = pre_bp2,
  post_bp = post_bp2,
  diff_bp = diff_bp2,
  post_under_120 = post_less_120_2
)
bp_df2

# print data frame nicely
knitr::kable(bp_df2)
```

2. Create and store a list with two elements:
  - 1st element named **treatment** and contains the first data frame
  - 2nd element named **placebo** and contains the second data frame

```{r}
# create treatment and placebo objects for each corresponding dataframe
treatment <- bp_df1
placebo <- bp_df2

# create list with treatment and placebo using list()
bp_exp <- list(treatment = treatment, placebo = placebo)
bp_exp
```

3. Access the first list element using three different types of syntax. 

```{r}
# first access using brackets [], returns a list containing the named element
bp_exp[1]

# second access using double brackets [[]], returns the actual data frame 
bp_exp[[1]]

# third access using $, also returns just the data frame itself
bp_exp$treatment
```

4. In one line, access the `placebo` data frame, `pre_bp2` column.

```{r}
# use [] and [[]] to access the pre_bp2 column as a data frame
bp_exp[[2]] [2:2]

# or use $ and [] to do the same 
bp_exp$placebo[2]

# or use just $ to return the vector only
bp_exp$placebo$pre_bp
```

