---
title: "Homework 2 - Member A"
author: "Alfonso Straffon"
format: html
editor: visual
---

## Task 1: Basic Vector Practice

### 1) Create two vectors using the `c()` function

The combine or `c()` function creates vectors by aggregating individual elements.
```{r}
# create vector with pre measurements using combine function
pre_bp <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)

# create vector with post measurements using combine function
post_bp <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)

```

### 2) Assign names to the vector elements using the `paste()` function

The `names()` function can be used to assign a character vector of names to the 
elements of a vector.
```{r}
# create names for the 20 subjects using a character vector
names(pre_bp) <- paste("Subject", 1:20, sep = "_")
names(post_bp) <- paste("Subject", 1:20, sep = "_")

pre_bp
post_bp

```

### 3) Calculate change in blood pressure

Here we are creating a new vector by subtracting two vectors of the same length.
```{r}
#subtract the post treatment level from pre treatment level for each subject and store as vector
diff_bp <- (pre_bp - post_bp)
diff_bp

```

This shows that two numeric vectors of equal length can be subtracted directly, and 
R performs the operation element by element. Of the 20 subjects, all but three showed 
decreases in blood pressure.

### 4) Calculate average change in blood pressure using `mean()` function

We use the `mean()` function to compute the average
```{r}
# 4) Calculate average
mean(diff_bp)
```

The average decrease in blood pressure was 17.

### 5) Identify which subjects showed a decrease in blood pressure

The comparison `diff_bp > 0` creates a logical vector, and the `which()` function 
returns the positions where that logical vector is `TRUE`.
```{r}
# we use a logical operator to determine positive values
which(diff_bp > 0)

```
In this case, the output gives the indices and names associated with positive changes.


### 6) Subset the vector of differences for only positive differences

```{r}
#one way of doing this is to create a position vector of positive differences
pos_indice <- which(diff_bp > 0)

# returns the elements of the positions listed in pos_indice
diff_bp[pos_indice]

```

We use `[]` in R to subset or select certain elements from an object.

### 7) Calculate average change in blood pressure for positive differences

```{r}
#use `mean()` function again
mean(diff_bp[pos_indice])
```

The average decrease in blood pressure was 20.65.

## Task 2: Basic Data Frame Practice

### 1) Create a data frame with four columns

```{r}
# create a separate character vector for the subject names
patient <- names(pre_bp)

#create data frame using `data.frame()` function
bp_df <- data.frame(patient, pre_bp, post_bp, diff_bp)
bp_df
```

The big teaching here is that the subject names, up until this step, were stored as the `names` attribute of the vectors and not as a separate column. In order to use them as the `patient` column in the data frame we first extracted them into their own character vector. We also note that vector names can automatically become data frame row names, but that row names are not the same thing as a data frame column.

### 2) Return data frame rows where `diff_bp` column is negative

```{r}

#from the data frame, keep rows where `diff_bp` is negative
bp_df[bp_df$diff_bp < 0, ]
```

This confirms that of the 20 subjects, three of them saw negative differences between the pre and post treatments.

### 3) Add a new column using a condition

```{r}
# identify in data frame where post_bp is less than 120 
bp_df$post_bp < 120 # this creates a logical vector

#create a new column with the logical vector
bp_df$less_120 <- bp_df$post_bp < 120
```

Worth noting that the comparison `bp_df$post_bp < 120`produced a logical vector, but 
it did not need to be stored as a separate vector object in order to assign it directly 
to the data frame column. Alternatively, we could have saved the logical vector as a 
separate vector object and then assigned it to the data frame, but this would have 
introduced an extra step.

### 4) Print the data frame in document

```{r}
#| warning:  false #this suppresses warning messages when rendering
#| message:  false #this suppresses package messages when rendering 

#print new data frame in document
knitr::kable(bp_df)
```

## Task 3: List Practice

### 1) Create a new data frame with placebo experiment

The code chunk that follows has most of the steps as in Task 1 and Task 2 above for 
placebo experiment.
```{r}
#| warning:  false #this suppresses warning messages when rendering
#| message:  false #this suppresses package messages when rendering 

#From Task 1
# 1) create vector with pre measurements using combine function
pre_bp_plc <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)

# 1) create vector with post measurements using combine function
post_bp_plc <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)

# 2) create names for the 10 subjects using a character vector
names(pre_bp_plc) <- paste("Subject", 1:10, sep = "_")
names(post_bp_plc) <- paste("Subject", 1:10, sep = "_")

pre_bp_plc
post_bp_plc

# 3) subtract the post treatment level from pre treatment level for each subject and store as vector
diff_bp_plc <- (pre_bp_plc - post_bp_plc)
diff_bp_plc

#From Task 2
# 1) create a separate character vector for the subject names
patient_plc <- names(pre_bp_plc)

# 1) create data frame using `data.frame()` function and use same column names as
# in treatment data frame
bp_df_plc <- data.frame(
  patient = patient_plc, 
  pre_bp = pre_bp_plc, 
  post_bp = post_bp_plc, 
  diff_bp = diff_bp_plc
)

bp_df_plc

# 3) create a new column with the logical vector
bp_df_plc$less_120 <- bp_df_plc$post_bp < 120

# 4) print new data frame in document
knitr::kable(bp_df_plc)

```

### 2) Create and store a list with two elements
It is worth noting a key difference between a vector and a list: A vector contains 
individual values of the same basic type, whereas a list can contain entire R objects 
as its elements, and those objects may or may not be the same type.

```{r}
#use the `list()` function to create a list and rename elements
bp_list <- list(treatment = bp_df, placebo = bp_df_plc)

#inspect structure of list_bp
str(bp_list)

```
### 3) Access the first list element using three different types of syntax

We covered three methods to access list elements.
```{r}
#single square brackets syntax
bp_list[1] # returns a list containing the first element

#double square brackets syntax 
bp_list[[1]] # returns the first element itself

#using the $ operator syntax
bp_list$treatment


```
### 4) Access the `placebo` data frame, `pre_bp` column

We can use the $ operator syntax in chain form to access a specific column in a 
given list and/or data frame.

```{r}
#Using a single line of code to access the pre_bp column in the `placebo` data frame
bp_list$placebo$pre_bp

```
## Task 4: Control Flow Practice

### 1) Creating an empty column in each of the data frames 

We create an empty column in the data frames called `status`. It is important to 
note the columns are being created in the list data frames, and do not carry backwards 
to the original data frames. 

```{r}
# create empty character status columns in both data frames stored in the list
bp_list$treatment$status <- character(20)
bp_list$placebo$status <- character(10)

```
### 2) Characterizing the post blood pressure measurements for the treatment study

First we establish the following criteria for the post blood pressure measurements:

- optimal (≤ 120)
- borderline (120 < bp ≤ 130)
- high (> 130)

We then use for loop and if/then/else logic to populate the `status` column
values in the treatment data frame.

```{r}
for (i in 1:20) {
  
  if (bp_list$treatment$post_bp[i] <= 120) {
    bp_list$treatment$status[i] = "optimal"
    
  } else if (bp_list$treatment$post_bp[i] <= 130) {
    bp_list$treatment$status[i] = "borderline"
    
  } else {bp_list$treatment$status[i] = "high"
    
  }
}

#inspect list data frame
bp_list$treatment
```

### 3) Characterizing the post blood pressure measurements for the placebo study

We repeat the above for loop and if/then/else logic step to populate the `status`
column for the placebo data frame.

```{r}
for (i in 1:10) {
  
  if (bp_list$placebo$post_bp[i] <= 120) {
    bp_list$placebo$status[i] = "optimal"
    
  } else if (bp_list$placebo$post_bp[i] <= 130) {
    bp_list$placebo$status[i] = "borderline"
    
  } else {
    bp_list$placebo$status[i] = "high"
  }
}

#inspect list data frame
bp_list$placebo
```
-END-
