---
title: "Homework 5"
author: "Alfonso Straffon"
format: html
---

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(httr)
library(jsonlite)
```

# Task 1 - Querying an API and a Tidy-Style Function

## Question 1

Use `GET()` from the `httr` package to return information about a topic of interest that has been in the news lately. Store the result as an R object.

```{r}
#this creates an HTTP response object
news_return <- GET("https://newsapi.org/v2/everything?q=prediction%20markets&from=2026-08-21&language=en&pageSize=50&apiKey=5666c93387864ff7b1e91fc2cef51c8a")

#to inspect structure
str(news_return, max.level = 1)
```

## Question 2

Parse what is returned and find the data frame containing the article information.

```{r}
#parsing converts raw JSON into a normal R object
#`jasonlite` is an R package used to read, parse, and convert JSON data into R objects
library(jsonlite)

#news_return$content -> grabs the raw API response
#rawToChar() -> converts the raw data into readable JSON text
#fromJSON() -> converts the JSON into an R object
#parsed <-  -> saves converted object
parsed <- fromJSON(rawToChar(news_return$content))
names(parsed)

#data frame containing article information only
str(parsed$articles)

```

## Question 3

Write a function that allows the user to query the API. The function should take:

- a title or subject to search for,
- a starting date,
- an API key.

Use the function twice to demonstrate that it works.

```{r}
# creating a function to run a query on NewsAPI
news_query <- function(subject, from_date, api_key) {
  # replaces blank spaces with proper URL spacing 
  subject <- gsub(" ", "%20", subject)
  # build URL with `paste0()` function, which joins text with no separator
  url <- paste0(
     "https://newsapi.org/v2/everything?q=", subject,
     "&from=", from_date,
     "&language=en&pageSize=50&apiKey=", api_key
  )
  
  # sends request to NewsAPI and stores the response
  news_return <- GET(url)
  
  # convert raw JSON response into an R object
  parsed <- fromJSON(rawToChar(news_return$content))
  
  # returns data frame containing the article information
  parsed$articles
}

#Test 1
test1 <- news_query(
  "prediction markets",
  "2026-09-18",
  "5666c93387864ff7b1e91fc2cef51c8a"
)

#Test 2
test2 <- news_query(
  "Clarity Act",
  "2026-09-18",
  "5666c93387864ff7b1e91fc2cef51c8a"
)
```

# Task 2 - Summarizing Roller Coaster Data

## Read in the Data

```{r}
coaster_data <- read_csv(
  "https://www4.stat.ncsu.edu/online/datasets/635USrollercoasters.csv"
)
```

## Question 4

Look at how the data is stored and determine whether everything makes sense. Convert variables to factors as needed.

```{r}
#checking structure of data to make sure it makes sense
str(coaster_data)

#converting data; note the ` ` for the column headers with special characters
coaster_data2 <- coaster_data |>
  mutate(State = as.factor(State),
         `Census Region` = as.factor(`Census Region`),
         AgeGroup = as.factor(AgeGroup),
         Scale = as.factor(Scale),
         Type = as.factor(Type),
         Design = as.factor(Design),
         `Inversions?` = as.factor(`Inversions?`))

str(coaster_data2)

```
In the original data, several categorical variables were stored as character variables, 
specifically `State`, `Census Region`, `AgeGroup`, `Scale`, `Type`, `Design`, and 
`Inversions?`. They were converted to *factors* using the `mutate()` and `across()` 
functions. As noted in class, factors are useful for categorical variables because 
they have a levels attribute that becomes helpful for ordering and plotting.

## Question 5

Document any missing values in the data.

```{r}

#from EDA material that shows us how to find missing data
colSums(is.na(coaster_data2))
```
Missing data are found in `Top Speed` = 74, `MaxHeight` = 8, `Drop` = 128, `Length` =33, `Duration` = 66, and `Make` = 36

# Categorical Variables

## Question 6

Create:

- a one-way contingency table,
- a two-way contingency table,
- a three-way contingency table.

Use `table()` and interpret one value from each resulting table.

### One-Way Table

```{r}
# One-way contingency table
table(coaster_data2$Type)

```
There are 635 roller coaster observations in the data. 530 are made from *steel* and
105 are made from *wood*.

### Two-Way Table

```{r}
# Two-way contingency table
table(coaster_data2$Scale, coaster_data2$Type)
```

The *scale* of the roller coaster further classifies the ride by its intensity level 
across the `type` variable. It shows that the *Extreme* intensity level is the most 
popular (216 + 53 = 269 rides) while the *Kiddie* intensity level is the least popular 
(27 + 1 = 28 rides).


### Three-Way Table

```{r}
# Three-way contingency table
table(
  coaster_data2$Scale,
  coaster_data2$Type,
  coaster_data2$`Inversions?`)
```

The `Inversions?` factor tells us the number of roller coasters that have loops. 
Of the 635 roller coasters in our sample there are 170,  or about 27%, that have 
loops. It is also evident that the material of choice to accommodate for loops is 
steel given it accounts for 166 of the 170 roller coasters.


## Question 7

Create a conditional two-way table using two different methods.

### Method 1 - Subset the Data First

```{r}
# Filter data to roller coasters with inversions
coaster_inversions <- coaster_data2 |>
  filter(`Inversions?` == "Yes")

# Create two-way conditional table
table(coaster_inversions$Scale, coaster_inversions$Type)
```

### Method 2 - Subset a Three-Way Table

```{r}
# Create three-way contingency table
coaster_table <- table(
  coaster_data2$Scale,
  coaster_data2$Type,
  coaster_data2$`Inversions?`
)

# Condition on `Inversions?` = Yes
coaster_table[, , "Yes"]
```

## Question 8

Create a two-way contingency table using `group_by()` and `summarize()`. Then use `pivot_wider()` so that the result resembles the output from `table()`.

```{r}

```

## Question 9

Create both a stacked bar graph and a side-by-side bar graph. Include appropriate axis labels and titles.

### Stacked Bar Graph

```{r}

```

### Side-by-Side Bar Graph

```{r}

```