---
title: "Homework 5"
author: "Maya Szymborski"
format: html
editor: source
---

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(httr)
```

# Task 1 - Querying an API and a Tidy-Style Function

## Question 1

Use `GET()` from the `httr` package to return information about a topic of interest that has been in the news lately. Store the result as an R object.

```{r}

```

## Question 2

Parse what is returned and find the data frame containing the article information.

```{r}

```

## Question 3

Write a function that allows the user to query the API. The function should take:

- a title or subject to search for,
- a starting date,
- an API key.

Use the function twice to demonstrate that it works.

```{r}

```

# Task 2 - Summarizing Roller Coaster Data

## Read in the Data

```{r}
coaster_data <- read_csv(
  "https://www4.stat.ncsu.edu/online/datasets/635USrollercoasters.csv"
)
```

## Question 4

Look at how the data is stored and determine whether everything makes sense. Convert variables to factors as needed.

```{r}

```

## Question 5

Document any missing values in the data.

```{r}

```

# Categorical Variables

## Question 6

Create:

- a one-way contingency table,
- a two-way contingency table,
- a three-way contingency table.

Use `table()` and interpret one value from each resulting table.

### One-Way Table

```{r}

```

**Interpretation:**

### Two-Way Table

```{r}

```

**Interpretation:**

### Three-Way Table

```{r}

```

**Interpretation:**

## Question 7

Create a conditional two-way table using two different methods.

### Method 1 - Subset the Data First

```{r}

```

### Method 2 - Subset a Three-Way Table

```{r}

```

## Question 8

Create a two-way contingency table using `group_by()` and `summarize()`. Then use `pivot_wider()` so that the result resembles the output from `table()`.

```{r}

```

## Question 9

Create both a stacked bar graph and a side-by-side bar graph. Include appropriate axis labels and titles.

### Stacked Bar Graph

```{r}

```

### Side-by-Side Bar Graph

```{r}

```
