---
title: "Homework 5"
author: "Alfonso Straffon"
format: html
---

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(httr)
library(jsonlite)
```

# Task 1 - Querying an API and a Tidy-Style Function

## Question 1

Use `GET()` from the `httr` package to return information about a topic of interest that has been in the news lately. Store the result as an R object.

```{r}
#this creates an HTTP response object
news_return <- GET("https://newsapi.org/v2/everything?q=prediction%20markets&from=2026-08-21&language=en&pageSize=50&apiKey=5666c93387864ff7b1e91fc2cef51c8a")

#to inspect structure
str(news_return, max.level = 1)
```

## Question 2

Parse what is returned and find the data frame containing the article information.

```{r}
#parsing converts raw JSON into a normal R object
#`jasonlite` is an R package used to read, parse, and convert JSON data into R objects
library(jsonlite)

#news_return$content -> grabs the raw API response
#rawToChar() -> converts the raw data into readable JSON text
#fromJSON() -> converts the JSON into an R object
#parsed <-  -> saves converted object
parsed <- fromJSON(rawToChar(news_return$content))
names(parsed)

#data frame containing article information only
str(parsed$articles)

```

## Question 3

Write a function that allows the user to query the API. The function should take:

- a title or subject to search for,
- a starting date,
- an API key.

Use the function twice to demonstrate that it works.

```{r}
# creating a function to run a query on NewsAPI
news_query <- function(subject, from_date, api_key) {
  # replaces blank spaces with proper URL spacing 
  subject <- gsub(" ", "%20", subject)
  # build URL with `paste0()` function, which joins text with no separator
  url <- paste0(
     "https://newsapi.org/v2/everything?q=", subject,
     "&from=", from_date,
     "&language=en&pageSize=50&apiKey=", api_key
  )
  
  # sends request to NewsAPI and stores the response
  news_return <- GET(url)
  
  # convert raw JSON response into an R object
  parsed <- fromJSON(rawToChar(news_return$content))
  
  # returns data frame containing the article information
  parsed$articles
}

#Test 1
test1 <- news_query(
  "prediction markets",
  "2026-09-18",
  "5666c93387864ff7b1e91fc2cef51c8a"
)

#Test 2
test2 <- news_query(
  "Clarity Act",
  "2026-09-18",
  "5666c93387864ff7b1e91fc2cef51c8a"
)
```

# Task 2 - Summarizing Roller Coaster Data

## Read in the Data

```{r}
coaster_data <- read_csv(
  "https://www4.stat.ncsu.edu/online/datasets/635USrollercoasters.csv"
)
```

## Question 4

Look at how the data is stored and determine whether everything makes sense. Convert variables to factors as needed.

```{r}

```

## Question 5

Document any missing values in the data.

```{r}

```

# Categorical Variables

## Question 6

Create:

- a one-way contingency table,
- a two-way contingency table,
- a three-way contingency table.

Use `table()` and interpret one value from each resulting table.

### One-Way Table

```{r}

```

**Interpretation:**


### Two-Way Table

```{r}

```

**Interpretation:**


### Three-Way Table

```{r}

```

**Interpretation:**


## Question 7

Create a conditional two-way table using two different methods.

### Method 1 - Subset the Data First

```{r}

```

### Method 2 - Subset a Three-Way Table

```{r}

```

## Question 8

Create a two-way contingency table using `group_by()` and `summarize()`. Then use `pivot_wider()` so that the result resembles the output from `table()`.

```{r}

```

## Question 9

Create both a stacked bar graph and a side-by-side bar graph. Include appropriate axis labels and titles.

### Stacked Bar Graph

```{r}

```

### Side-by-Side Bar Graph

```{r}

```