---
title: "Homework 5"
author: "Maya Szymborski"
format: html
editor: source
---

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(httr)
library(jsonlite)
```

# Task 1 - Querying an API and a Tidy-Style Function

## Question 1

Use `GET()` from the `httr` package to return information about a topic of interest that has been in the news lately. Store the result as an R object.

```{r}
# use GET() to send a GET request to the news API 

new_species <- GET("https://newsapi.org/v2/everything?q=Leopardus%20tilcayo&from=2026-9-18&language=en&pageSize=50&apiKey=ef59b58fce7542e0b4b17976d880650a")
```

## Question 2

Parse what is returned and find the data frame containing the article information.

```{r}
# convert from raw data to text (`rawToChar()`) 
# and parse the JSON text into an R object (`fromJSON()`)
parsed <- fromJSON(rawToChar(new_species$content))

# convert articles into a tibble
article_info <- as_tibble(parsed$articles)

# display names of variables 
names(article_info)

# display relevant information
article_info |>
  select(author, title, url)
```

## Question 3

Write a function that allows the user to query the API. The function should take:

- a title or subject to search for,
- a starting date,
- an API key.

Use the function twice to demonstrate that it works.

```{r}
# create a function that queries the news API

query_news_API <- function(subject, start_date, your_API_key) {
  
  # query news API
  response <- GET("https://newsapi.org/v2/everything",
                  query = list(
                    q = subject,
                    from = start_date,
                    language = "en",
                    pageSize = 50,
                    apiKey = your_API_key
                  ))
  # no need to build URL because of `query = list(...)`
  
  # convert & parse response
  parsed <- fromJSON(rawToChar(response$content))
  
  # convert articles to tibble
  article_info <- as_tibble(parsed$articles)
  
  return(article_info)
}

# first function use
species_news <- query_news_API(
  subject = "Leopardus tilcayo",
  start_date = "2026-09-18",
  your_API_key = "ef59b58fce7542e0b4b17976d880650a"
)
species_news

# second function use
ESA_news <- query_news_API(
  subject = "Endangered Species Act Trump",
  start_date = "2026-09-18",
  your_API_key = "ef59b58fce7542e0b4b17976d880650a"
)
ESA_news
```

can also do
```{r}
query_news_API_2 <- function(subject, start_date, your_API_key) {
  
  # create the API URL
  url <- paste0("https://newsapi.org/v2/everything?q=",
                     subject,
                     "&from=",
                     start_date,
                     "&language=en&pageSize=50&apiKey=",
                     your_API_key,
                     sep = ""
  )
  
  # send GET request 
  reponse <- GET(url)
  
  # convert & parse reponse 
  parsed_data <- fromJSON(rawToChar(response$content))
  
  # convert to tibble
  articles <- as_tibble(parsed_data$articles)
  
  return(articles)
  
```

# Task 2 - Summarizing Roller Coaster Data

## Read in the Data

```{r}
coaster_data <- read_csv(
  "https://www4.stat.ncsu.edu/online/datasets/635USrollercoasters.csv"
)
head(coaster_data)
```

## Question 4

Look at how the data is stored and determine whether everything makes sense. Convert variables to factors as needed.

```{r}
# look at structure of data
str(coaster_data) # should `YearOpened` be  chr instead of num? 

# turn applicable variables into factors (variables that  represent categories)
coaster_data <- coaster_data |>
  mutate(State = as.factor(State),
         `Census Region` = as.factor(`Census Region`),
         AgeGroup = as.factor(AgeGroup),
         Scale = as.factor(Scale),
         Type = as.factor(Type),
         Design = as.factor(Design),
         `Inversions?` = as.factor(`Inversions?`)
  )

str(coaster_data)
```

## Question 5

Document any missing values in the data.

```{r}
# check for missing values using `is.na()`

colSums(is.na(coaster_data))
```

`TopSpeed`, `MaxHeight`, `Drop`, `Length`, `Duration`, and `Make` have missing values

# Categorical Variables

## Question 6

Create:

- a one-way contingency table,
- a two-way contingency table,
- a three-way contingency table.

Use `table()` and interpret one value from each resulting table.

### One-Way Table

```{r}
# create a one-way contingency table

table(coaster_data$State)
```

**Interpretation:**
There are 15 roller coasters from North Carolina in this dataset. 

### Two-Way Table

```{r}
# create a two-way contingency table

table(coaster_data$State, coaster_data$Type)
```

**Interpretation:**
Of the 15 roller coasters from NC, 14 are Steel and only one is Wood. 

### Three-Way Table

```{r}
# create a three-way contingency table

table(coaster_data$State, coaster_data$Scale, coaster_data$Type)
```

**Interpretation:**
Surprisingly, the one wooden roller coaster in NC is classified as an Extreme roller coaster (I don't think I would trust going on it). The rest of the coasters are Steel, with 8 coasters classified as Extreme, 2coasters  as Family, 2 classified as Kiddie, and 2 classified as Thrill.

## Question 7

Create a conditional two-way table using two different methods.

### Method 1 - Subset the Data First

```{r}
# check how State values are coded
unique(coaster_data$State)

# use `filter()` to subset the data to roller coasters only in North Carolina

NC_coasters <- coaster_data |>
  filter(State == "North Carolina")

# check
NC_coasters

# create a two-way conditional table comparing coaster type and age group
table(NC_coasters$Type, NC_coasters$AgeGroup)
```
- The 1 Wood roller coaster in NC is classified as Recent, so maybe I trust it. But how recent? 

```{r}
# check the opening year of the wooden coaster in NC
NC_coasters |>
  filter(Type == "Wood") |>
  select(Coaster, Park, `Year Opened`, AgeGroup)
```
- The 1 Wood roller coaster, called the Hurler, was opened in 1994 at Carowinds. I think I would rather ride a newer wooden roller coaster, if I rode a wooden roller coaster at all. 

- Haha, got a little sidetracked here!

### Method 2 - Subset a Three-Way Table

```{r}
# create a three-way contigency table
NC_coasters2 <- table(
  coaster_data$Type, 
  coaster_data$AgeGroup, 
  coaster_data$State
)

# condition is on State
NC_coasters2[, , "North Carolina"]
```

## Question 8

Create a two-way contingency table using `group_by()` and `summarize()`. Then use `pivot_wider()` so that the result resembles the output from `table()`.

```{r}
# create the same two-way contingency table as before 
# but by using `group_by()` and `summarize()`

coaster_data |>
  group_by(State, Type) |>
  summarize(count = n())

# use `pivot_wider()` so result resembles the output from `table()`

coaster_data |>
  group_by(State, Type) |>
  summarize(count = n()) |>
  pivot_wider(names_from = Type, 
              values_from = count)
```

## Question 9

Create both a stacked bar graph and a side-by-side bar graph. Include appropriate axis labels and titles.

### Stacked Bar Graph

```{r}
library(ggplot2)
help(ggplot) # to see syntax; usually would do in console (or can look at cheat sheet)

# create a simple stacked bar graph
ggplot(
  data = coaster_data, 
  aes(x = AgeGroup, fill = Type)
  ) + 
  geom_bar() + 
  labs(
    x = "Roller Coaster Age Group", 
    y = "Number of Roller Coasters", 
    title = "Roller Coaster Type by Age Group")
```


### Side-by-Side Bar Graph

```{r}
# create a simple side-by-side bar graph 
# use `geom_bar(position = "dodge")` (from help(geom_bar))

ggplot(
  data = coaster_data, 
  aes(x = AgeGroup, fill = Type)
  ) + 
  geom_bar(position = "dodge") + 
  labs(
    x = "Roller Coaster Age Group", 
    y = "Number of Roller Coasters", 
    title = "Roller Coaster Type by Age Group")
```
