---
title: "Homework 5"
author: " Daria Smyslova, Maya Szymborski, Michael Costello & Alfonso Straffon"
format: html
---

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(httr)
library(jsonlite)
```

# Task 1 - Querying an API and a Tidy-Style Function

## Question 1

Use `GET()` from the `httr` package to return information about a topic of interest that has been in the news lately. Store the result as an R object.

```{r}
#this creates an HTTP response object
news_return <- GET("https://newsapi.org/v2/everything?q=prediction%20markets&from=2026-08-23&language=en&pageSize=50&apiKey=5666c93387864ff7b1e91fc2cef51c8a")

#to inspect structure
str(news_return, max.level = 1)
```

## Question 2

Parse what is returned and find the data frame containing the article information.

```{r}
# Parse the API response
parsed <- fromJSON(rawToChar(news_return$content))

# Convert the article information to a tibble
article_info <- as_tibble(parsed$articles)

glimpse(article_info)

# Convert source back to a list column
article_info <- article_info |>
  mutate(
    source = pmap(source, \(...) list(...))
  )

glimpse(article_info)

```
NOTE: Using `fromJSON()` automatically simplifies the nested structure into a *data frame*, 
so we have to convert the `source` column back to a *list* as requested in the assignment.


## Question 3

Write a function that allows the user to query the API. The function should take:

- a title or subject to search for,
- a starting date,
- an API key.

Use the function twice to demonstrate that it works.

```{r}
news_query <- function(subject, start_date, api_key) {
  
  # Send request to NewsAPI
  response <- GET(
    "https://newsapi.org/v2/everything",
    query = list(
      q = subject,
      from = start_date,
      language = "en",
      pageSize = 50,
      apiKey = api_key
    )
  )
  
  # Parse the API response
  parsed <- fromJSON(rawToChar(response$content))

  # Convert the article information into a tibble
  article_info <- as_tibble(parsed$articles)
  
  # Convert source back to a list column
  article_info <- article_info |>
    mutate(
      source = pmap(source, \(...) list(...))
    )

  return(article_info)
}

#Test 1
test1 <- news_query(
  "prediction markets",
  "2026-09-18",
  "5666c93387864ff7b1e91fc2cef51c8a"
)

glimpse(test1) #before
head(test1$title)

#Test 2
test2 <- news_query(
  "Clarity Act",
  "2026-09-18",
  "5666c93387864ff7b1e91fc2cef51c8a"
)
glimpse(test2) #after
head(test2$title)


```

# Task 2 - Summarizing Roller Coaster Data

## Read in the Data

```{r}
coaster_data <- read_csv(
  "https://www4.stat.ncsu.edu/online/datasets/635USrollercoasters.csv"
)
```

## Question 4

Look at how the data is stored and determine whether everything makes sense. Convert variables to factors as needed.

```{r}
#checking structure of data to make sure it makes sense
str(coaster_data)

#converting data; note the ` ` for the column headers with special characters
coaster_data2 <- coaster_data |>
  mutate(State = as.factor(State),
         `Census Region` = as.factor(`Census Region`),
         AgeGroup = as.factor(AgeGroup),
         Scale = as.factor(Scale),
         Type = as.factor(Type),
         Design = as.factor(Design),
         `Inversions?` = as.factor(`Inversions?`),
         Make = as.factor(Make)) #added make as factor

str(coaster_data2)

```
In the original data, several categorical variables were stored as character variables, 
specifically `State`, `Census Region`, `AgeGroup`, `Scale`, `Type`, `Design`, `Inversions?`, 
and `Make`. They were converted to *factors* using the `mutate()` function. As noted 
in class, factors are useful for categorical variables because they have a 
levels attribute that becomes helpful for ordering and plotting.

## Question 5

Document any missing values in the data.

```{r}

#from EDA material that shows us how to find missing data
colSums(is.na(coaster_data2))
```
Missing data are found in `TopSpeed` = 74, `MaxHeight` = 8, `Drop` = 128, `Length` =33, `Duration` = 66, and `Make` = 36

# Categorical Variables

## Question 6

Create:

- a one-way contingency table,
- a two-way contingency table,
- a three-way contingency table.

Use `table()` and interpret one value from each resulting table.

### One-Way Table

```{r}
# One-way contingency table
table(coaster_data2$Type)

```
There are 635 roller coaster observations in the data. 530 are made from *steel* and
105 are made from *wood*.



### Two-Way Table

```{r}
# Two-way contingency table
table(coaster_data2$Scale, coaster_data2$Type)
```

The *scale* of the roller coaster further classifies the ride by its intensity level 
across the `type` variable. It shows that the *Extreme* intensity level is the most 
popular (216 + 53 = 269 rides) while the *Kiddie* intensity level is the least popular 
(27 + 1 = 28 rides).


### Three-Way Table

```{r}
# Three-way contingency table
table(
  coaster_data2$Scale,
  coaster_data2$Type,
  coaster_data2$`Inversions?`)
```

The `Inversions?` factor tells us the number of roller coasters that have inversions. 
Of the 635 roller coasters in our sample there are 170,  or about 27%, that have 
loops. It is also evident that the material of choice to accommodate for loops is 
steel given it accounts for 166 of the 170 roller coasters.


## Question 7

Create a conditional two-way table using two different methods.

### Method 1 - Subset the Data First

```{r}
# Filter data to roller coasters with inversions
coaster_inversions <- coaster_data2 |>
  filter(`Inversions?` == "Yes")

# Create two-way conditional table
table(coaster_inversions$Scale, coaster_inversions$Type)
```

### Method 2 - Subset a Three-Way Table

```{r}
# Create three-way contingency table
coaster_table <- table(
  coaster_data2$Scale,
  coaster_data2$Type,
  coaster_data2$`Inversions?`
)

# Condition on `Inversions?` = Yes
coaster_table[, , "Yes"]
```

## Question 8

Create a two-way contingency table using `group_by()` and `summarize()`. Then use `pivot_wider()` so that the result resembles the output from `table()`.

```{r}
#| message: false

coaster_data2 |>
  group_by(Scale, Type) |>
  summarize(count = n()) |>
  pivot_wider(
    names_from = Type, #determines column headings, in this case `Type`
    values_from = count
  )
```

## Question 9

Create both a stacked bar graph and a side-by-side bar graph. Include appropriate axis labels and titles.

### Stacked Bar Graph

```{r}

ggplot(
  data = coaster_data2,
  aes(x = Scale, fill = Type)
) +
  geom_bar() +
  labs(
    x = "Roller Coaster Scale",
    y = "Number of Roller Coasters",
    fill = "Coaster Type",
    title = "Roller Coaster Type by Scale"
  )
```

### Side-by-Side Bar Graph

```{r}

ggplot(
  data = coaster_data2,
  aes(x = Scale, fill = Type)
) +
  geom_bar(position = "dodge") +
  labs(
    x = "Roller Coaster Scale",
    y = "Number of Roller Coasters",
    fill = "Coaster Type",
    title = "Roller Coaster Type by Scale"
  )
```
