---
title: "Homework 5"
author: "Maya Szymborski"
format: html
editor: source
---

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(httr)
```

# Task 1 - Querying an API and a Tidy-Style Function

## Question 1

Use `GET()` from the `httr` package to return information about a topic of interest that has been in the news lately. Store the result as an R object.

```{r}

```

## Question 2

Parse what is returned and find the data frame containing the article information.

```{r}

```

## Question 3

Write a function that allows the user to query the API. The function should take:

- a title or subject to search for,
- a starting date,
- an API key.

Use the function twice to demonstrate that it works.

```{r}

```

# Task 2 - Summarizing Roller Coaster Data

## Read in the Data

```{r}
coaster_data <- read_csv(
  "https://www4.stat.ncsu.edu/online/datasets/635USrollercoasters.csv"
)
head(coaster_data)
```

## Question 4

Look at how the data is stored and determine whether everything makes sense. Convert variables to factors as needed.

```{r}
# look at structure of data
str(coaster_data) # should `YearOpened` be  chr instead of num? 

# turn applicable variables into factors (variables that  represent categories)
coaster_data <- coaster_data |>
  mutate(State = as.factor(State),
         `Census Region` = as.factor(`Census Region`),
         AgeGroup = as.factor(AgeGroup),
         Scale = as.factor(Scale),
         Type = as.factor(Type),
         Design = as.factor(Design),
         `Inversions?` = as.factor(`Inversions?`)
  )

str(coaster_data)
```

## Question 5

Document any missing values in the data.

```{r}
# check for missing values using `is.na()`

colSums(is.na(coaster_data))
```

`TopSpeed`, `MaxHeight`, `Drop`, `Length`, `Duration`, and `Make` have missing values

# Categorical Variables

## Question 6

Create:

- a one-way contingency table,
- a two-way contingency table,
- a three-way contingency table.

Use `table()` and interpret one value from each resulting table.

### One-Way Table

```{r}
# create a one-way contingency table

table(coaster_data$State)
```

**Interpretation:**
There are 15 roller coasters from North Carolina in this dataset. 

### Two-Way Table

```{r}
# create a two-way contingency table

table(coaster_data$State, coaster_data$Type)
```

**Interpretation:**
Of the 15 roller coasters from NC, 14 are Steel and only one is Wood. 

### Three-Way Table

```{r}
# create a three-way contingency table

table(coaster_data$State, coaster_data$Scale, coaster_data$Type)
```

**Interpretation:**
Surprisingly, the one wooden roller coaster in NC is classified as an Extreme roller coaster (I don't think I would trust going on it). The rest of the coasters are Steel, with 8 coasters classified as Extreme, 2coasters  as Family, 2 classified as Kiddie, and 2 classified as Thrill.

## Question 7

Create a conditional two-way table using two different methods.

### Method 1 - Subset the Data First

```{r}
# check how State values are coded
unique(coaster_data$State)

# use `filter()` to subset the data to roller coasters only in North Carolina

NC_coasters <- coaster_data |>
  filter(State == "North Carolina")

# check
NC_coasters

# create a two-way conditional table comparing coaster type and age group
table(NC_coasters$Type, NC_coasters$AgeGroup)
```
- The 1 Wood roller coaster in NC is classified as Recent, so maybe I trust it. But how recent? 

```{r}
# check the opening year of the wooden coaster in NC
NC_coasters |>
  filter(Type == "Wood") |>
  select(Coaster, Park, `Year Opened`, AgeGroup)
```
- The 1 Wood roller coaster, called the Hurler, was opened in 1994 at Carowinds. I think I would rather ride a newer wooden roller coaster, if I rode a wooden roller coaster at all. 

- Haha, got a little sidetracked here!

### Method 2 - Subset a Three-Way Table

```{r}
# create a three-way contigency table
NC_coasters2 <- table(
  coaster_data$Type, 
  coaster_data$AgeGroup, 
  coaster_data$State
)

# condition is on State
NC_coasters2[, , "North Carolina"]
```

## Question 8

Create a two-way contingency table using `group_by()` and `summarize()`. Then use `pivot_wider()` so that the result resembles the output from `table()`.

```{r}

```

## Question 9

Create both a stacked bar graph and a side-by-side bar graph. Include appropriate axis labels and titles.

### Stacked Bar Graph

```{r}

```

### Side-by-Side Bar Graph

```{r}

```
