---
title: "Homework 4"
author: "Your Name"
format: html
editor: source
---

### Task 1: Conceptual Questions

#### 1. Rewrite the following nested function call using baseR’s chaining operator:

```{r}
library(tidyverse)
arrange(filter(select(as_tibble(iris), starts_with("Petal"), Species), Petal.Length < 1.55), Species)
```

#### 2. What is meant by long format data and wide format data? Which do we generally prefer for statistical analysis?

### Task 2:

*Note: Use chaining where possible!*

#### 3. **Glass data:** In HW 3 you read in the glass data from https://www4.stat.ncsu.edu/~online/datasets/glass.data. Let’s reread that data (use your code from HW 3) that also names the columns. Recall the last variable being `Type of Glass` with values of:– 1 building_windows_float_processed, – 2 building_windows_non_float_processed, – 3 vehicle_windows_float_processed, – 4 vehicle_windows_non_float_processed (none in this database), – 5 containers, – 6 tableware, – 7 headlamps. Start a chain on your data object that overwrites the `Type_of_glass` variable using `mutate()`.

• Create a `factor` version of the variable that replaces the numbers with string labels (that is, replace 1 with “building_windows_float_processed”, 2 with “building_win. . . ”, etc. - see the variable descriptions above to give meaningful values). 

• This could also be done by first using an `ifelse()` type call and then converting to a factor. Either way is fine!

#### 4. Continue your chain and keep only observations where the `Fe` variable is less than 0.2 and the `Type of Glass` is either “tableware” or “headlamp”.

```{r}
library(tidyverse)
glass_data <- read_csv("https://www4.stat.ncsu.edu/~online/datasets/glass.data",   col_names = 
  c("Id", "RI", "Na", "Mg", "Al", "Si", "K", "Ca", "Ba", "Fe", "TypeOfGlass")
)
```

#### 5. **Yeast data** In HW 3 you read in the Yeast data from https://www4.stat.ncsu.edu/~online/datasets/yeast.data. Read this data into `R` directly from the URL, adding column names as well. Start a chain that removes the `seq_name` and `nuc` columns.

#### 6. Continue your chain to add columns corresponding to the mean and median of each numeric variable (`mcg`, `gvh`, `alm`, `mit`, `erl`, `pox`, and `vac`) at each `class` grouping (see the `across()` function as we did in the `dplyr` video!).

```{r}
yeast_data <- read_table("https://www4.stat.ncsu.edu/~online/datasets/yeast.data", col_names = c("seq_name", "mcg", "gvh", "alm", "mit", "erl", "pox", "vac", "nuc", "class"))
```


#### 7. **Wine Data** In HW 3 you read in two data sets on wine. Read those two in and combine them as was done there. Start a chain on your new combined data object to filter the data to only look at high-quality wines (`quality` > 6.5) and wines that have a reasonable alcohol value (`alcohol` < 13.2).

```{r}
#Read in White wine data
library(readxl)
wine_data <- read_excel("data/white-wine.xlsx")
#Add in type = White
wine_data <- wine_data |> mutate(type = "white")

#Read in the Red wine data
wine_data2 <- 
  read_delim("https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv", 
  delim = ";"
)
#Add in type = Red
wine_data2 <- wine_data2 |> mutate(type = "red")

#Combine the two dataset together
wine_data_final <- dplyr::bind_rows(wine_data, wine_data2)

#Read in and apply the Wine Labels
wine_data_lab <- read_excel("data/white-wine.xlsx", sheet =2)
colnames(wine_data_final) <- wine_data_lab[[1]]
wine_data_final
```

#### 8. Continue your chain to now sort the data from highest quality to lowest.

#### 9. Continue your chain to select only the variables that contain `acid`, the `alcohol` variable, the `type` variable, and the `quality` variable.

#### 10. Continue your chain to add the mean and standard deviation of the `alcohol` variable to the data set for each setting of the `quality` variable.

```{r}

```

### Task 3: Database Practice

#### 11. Download the `Lahman.db` file associated with the *Connecting to Databases* notes/video. Connect to the database and then look at all of the tables in the database.

```{r}

```

#### 12. Use the `tbl()` function and `dplyr` to **return** all of the data from the `Teams` table for the year 2015.

```{r}

```

#### 13. Repeat the above by using the `sql()` function within `tbl()` (here you have to write actual SQL!).
* Note: Either use dplyr or write SQL queries for the rest of these questions!*

```{r}

```

#### 14. Return all of the players in the hall of fame, the year they were voted into the hall of fame, and their category (only those three variables!). See the `HallOfFame` table, the inducted variable is important here.

```{r}

```

#### 15. Combine the table from the previous question with the `People` table in order to have the `nameFirst` and `nameLast` names added to what was returned above.

```{r}

```

#### 16. Return only the `playerID` (manager ID in this case), `G`, `W`, and `L` columns from the `Managers` table. Use chaining to then: 

• Determine the overall win/loss records (sum of wins and sum of losses) for each of these hall of fame managers. – We haven’t quite covered this so I’ll help out. We want to use `group_by(`) with `summarize()` rather than `mutate()`. Something like this:

```{r}
tbl(con, "Managers") |>
select(...) |>
group_by(playerID) |>
summarize(G_managed = sum(G, na.rm = TRUE),
		Total_W = sum(W, na.rm = TRUE),
		Total_L = sum(L, na.rm = TRUE))
```

• Create a new variable that is the career win/loss percentage (`Total_W/G_managed`). (Note: I did this after `collect()` otherwise the column type has to be set. . . ) 

• Sort the resulting data by the win/loss percentage variable (from largest to smallest).

```{r}

```

#### 17. Use the results from the previous two questions to answer this one! Return information for only the people that managed a team and were inducted into the hall of fame (regardless of their `category` in the hall of fame - you want to return all people in the hall of fame that ever managed at all). (Just return all of the variables from both of the above two questions - 9 total columns)

```{r}

```

