---
title: "Homework 4"
author: "Alfonso Straffon"
format: html
editor: source
---

### Task 1: Conceptual Questions

#### 1. Rewrite the following nested function call using baseR’s chaining operator:

```{r}

library(tidyverse)
arrange(filter(select(as_tibble(iris), starts_with("Petal"), Species), Petal.Length < 1.55), Species)

#inspect `iris` raw data
iris
class(iris)

#using baseR chaining operator to recreate the nested function
iris |>
  as_tibble() |>
  select(starts_with("Petal"), Species) |>
  filter(Petal.Length < 1.55) |>
  arrange(Species)

```
So basically I am telling R to 1) look for the `iris` dataframe, 2) convert it to 
a tibble, 3) select the column(s) that start with "Petal" AND the Species column, 
4) apply a filter to the rows using Petal.Length column condition of less than 1.55, 
and 5) arrange in order by `Species`in alphabetical order.


#### 2. What is meant by long format data and wide format data? Which do we generally prefer for statistical analysis?
Long format data basically has one observation per row, while wide format data can 
have multiple observations across different columns in the same row. For statistical 
analysis working with long format is generally easier for when plotting, summarizing, 
and modeling data.

### Task 2:

*Note: Use chaining where possible!*

#### 3. **Glass data:** In HW 3 you read in the glass data from https://www4.stat.ncsu.edu/\~online/datasets/glass.data. Let’s reread that data (use your code from HW 3) that also names the columns. Recall the last variable being `Type of Glass` with values of:– 1 building_windows_float_processed, – 2 building_windows_non_float_processed, – 3 vehicle_windows_float_processed, – 4 vehicle_windows_non_float_processed (none in this database), – 5 containers, – 6 tableware, – 7 headlamps. Start a chain on your data object that overwrites the `Type_of_glass` variable using `mutate()`.

```{r}

#From the `tidyverse` we use the `readr` package and the `read_csv()` function
#to read in the data. We also use the `col_names` argument to name the columns.

glass <- readr::read_csv(
  "https://www4.stat.ncsu.edu/~online/datasets/glass.data",
  col_names = c("Id", "RI", "Na", "Mg", "Al", "Si", "K", "Ca", "Ba", "Fe", "Type_of_glass")
)

#prints tibble
glass
```

• Create a `factor` version of the variable that replaces the numbers with string labels (that is, replace 1 with “building_windows_float_processed”, 2 with “building_win. . . ”, etc. - see the variable descriptions above to give meaningful values).

```{r}
glass |>
  mutate(
    Type_of_glass = factor(
      Type_of_glass,
      levels = c(1, 2, 3, 4, 5, 6, 7),
      labels = c(
        "building_windows_float_processed",
        "building_windows_non_float_processed",
        "vehicle_windows_float_processed",
        "vehicle_windows_non_float_processed",
        "containers",
        "tableware",
        "headlamps"
      )
    )
  )
```

• This could also be done by first using an `ifelse()` type call and then converting to a factor. Either way is fine!

```{r}
#ALTERNATIVE APPROACH
#using if else seems to be more complicated so will stick with first approach
glass |>
  mutate(Type_of_glass = factor(
    ifelse(
    Type_of_glass == 1,
    "building_windows_float_processed",
    ifelse(
    Type_of_glass == 2,
    "building_windows_non_float_processed",
    ifelse(
    Type_of_glass == 3,
    "vehicle_windows_float_processed",
    ifelse(
    Type_of_glass == 4,
    "vehicle_windows_non_float_processed",
    ifelse(
    Type_of_glass == 5,
    "containers",
    ifelse(
    Type_of_glass == 6,
    "tableware",
    ifelse(
    Type_of_glass == 7,
    "headlamps",
    NA)))))))))
```

#### 4. Continue your chain and keep only observations where the `Fe` variable is less than 0.2 and the `Type of Glass` is either “tableware” or “headlamp”.

```{r}
#using baseR chaining operator 

glass |>
  mutate(
    Type_of_glass = factor(
      Type_of_glass,
      levels = c(1, 2, 3, 4, 5, 6, 7),
      labels = c(
        "building_windows_float_processed",
        "building_windows_non_float_processed",
        "vehicle_windows_float_processed",
        "vehicle_windows_non_float_processed",
        "containers",
        "tableware",
        "headlamps"
      )
    )
  ) |>
  filter(
    Fe < 0.2,
    Type_of_glass == "tableware" |
      Type_of_glass == "headlamps"
  )

```

#### 5. **Yeast data** In HW 3 you read in the Yeast data from https://www4.stat.ncsu.edu/\~online/datasets/yeast.data. Read this data into `R` directly from the URL, adding column names as well. Start a chain that removes the `seq_name` and `nuc` columns.

```{r}
#From the `tidyverse` we use the `readr` package and the `read_table()` function
#to read in the data. We also use the `col_names` argument to name the columns.
yeast <- readr::read_table(
  "https://www4.stat.ncsu.edu/~online/datasets/yeast.data",
  col_names = c("seq_name", "mcg", "gvh", "alm", "mit", "erl", "pox", "vac", "nuc", "class")
)

#prints tibble
yeast

#start chain to remove `seq_name` and `nuc` columns
yeast |>
  select(-seq_name, -nuc)

```
#### 6. Continue your chain to add columns corresponding to the mean and median of each numeric variable (`mcg`, `gvh`, `alm`, `mit`, `erl`, `pox`, and `vac`) at each `class` grouping (see the `across()` function as we did in the `dplyr` video!).

```{r}
yeast |>
  select(-seq_name, -nuc) |>
  group_by(class) |>
  mutate(
    across( #note: `across()` combines original column name + "_" + function name
      mcg:vac,
      list(mean = mean, median = median)
    )
  )
```

#### 7. **Wine Data** In HW 3 you read in two data sets on wine. Read those two in and combine them as was done there. Start a chain on your new combined data object to filter the data to only look at high-quality wines (`quality` \> 6.5) and wines that have a reasonable alcohol value (`alcohol` \< 13.2).

```{r}
# Read in white wine data
white_wine <- readxl::read_excel("data/white-wine.xlsx", sheet = 1)

# Read in alternate variable names
white_wine_2 <- readxl::read_excel("data/white-wine.xlsx", sheet = 2)

# Replace white wine column names
colnames(white_wine) <- white_wine_2[[1]]

# Add wine type
white_wine$type <- "white"

# Read in red wine data
red_wine <- readr::read_delim(
  "https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv",
  delim = ";"
)

# Replace red wine column names
colnames(red_wine) <- white_wine_2[[1]]

# Add wine type
red_wine$type <- "red"

# Combine the two data sets
wine <- dplyr::bind_rows(white_wine, red_wine)


#now we filter to high-quality wines with alcohol less than 13.2
wine |>
  filter(
    quality > 6.5,
    alcohol < 13.2
  )
```

#### 8. Continue your chain to now sort the data from highest quality to lowest.

```{r}
wine |>
  filter(
    quality > 6.5,
    alcohol < 13.2
  ) |>
  arrange(desc(quality))
```


#### 9. Continue your chain to select only the variables that contain `acid`, the `alcohol` variable, the `type` variable, and the `quality` variable.
```{r}
wine |>
  filter(
    quality > 6.5,
    alcohol < 13.2
  ) |>
  arrange(desc(quality)) |>
  select(
    contains("acid"),
    alcohol,
    type,
    quality
  )
```

#### 10. Continue your chain to add the mean and standard deviation of the `alcohol` variable to the data set for each setting of the `quality` variable.

```{r}
wine |>
  filter(
    quality > 6.5,
    alcohol < 13.2
  ) |>
  arrange(desc(quality)) |>
  select(
    contains("acid"),
    alcohol,
    type,
    quality
  ) |>
  group_by(quality) |>
  mutate(
    mean_alcohol = mean(alcohol), #unlike the yeast example we don't use `across()'
    sd_alcohol = sd(alcohol)
  )
```

### Task 3: Database Practice

#### 11. Download the `Lahman.db` file associated with the *Connecting to Databases* notes/video. Connect to the database and then look at all of the tables in the database.

```{r}
#this loads the database interface package
library(DBI)

#we create an object to store the database connection
con <- dbConnect(
  RSQLite::SQLite(),
  "data/lahman.db"
)

# to look at all the tables in the database 
dbListTables(con)
```

#### 12. Use the `tbl()` function and `dplyr` to **return** all of the data from the `Teams` table for the year 2015.

```{r}
library(dplyr)

tbl(con, "Teams") |>
  filter(yearID == 2015)

```

#### 13. Repeat the above by using the `sql()` function within `tbl()` (here you have to write actual SQL!).

- Note: Either use dplyr or write SQL queries for the rest of these questions!\*

```{r}
tbl(
  con,
  sql("SELECT * FROM Teams WHERE yearID = 2015")
)
```

#### 14. Return all of the players in the hall of fame, the year they were voted into the hall of fame, and their category (only those three variables!). See the `HallOfFame` table, the inducted variable is important here.

```{r}
#dplyr version
tbl(con, "HallOfFame") |>
  filter(inducted == "Y") |>
  select(playerID, yearID, category)

#alternative SQL version
tbl(
  con,
  sql("
    SELECT playerID, yearID, category
    FROM HallOfFame
    WHERE inducted = 'Y'
  ")
)
```

#### 15. Combine the table from the previous question with the `People` table in order to have the `nameFirst` and `nameLast` names added to what was returned above.

```{r}
# dplyr version
# will create an object first
hall_of_fame <- tbl(con, "HallOfFame") |>
  filter(inducted == "Y") |>
  select(playerID, yearID, category)

people <- tbl(con, "People") |>
  select(playerID, nameFirst, nameLast)

#join tables
hall_of_fame_names <- hall_of_fame |>
  left_join(people, by = "playerID") |>
  collect()

#inspect
hall_of_fame_names
```

#### 16. Return only the `playerID` (manager ID in this case), `G`, `W`, and `L` columns from the `Managers` table. Use chaining to then:

• Determine the overall win/loss records (sum of wins and sum of losses) for each of these hall of fame managers. – We haven’t quite covered this so I’ll help out. We want to use `group_by(`) with `summarize()` rather than `mutate()`. Something like this:

```{r}
# will create an object first
managers <- tbl(con, "Managers") |>
  select(playerID, G, W, L) |>
  group_by(playerID) |>
  summarize(
    G_managed = sum(G, na.rm = TRUE),
    Total_W = sum(W, na.rm = TRUE),
    Total_L = sum(L, na.rm = TRUE)
  )
#inspect object
managers
```

• Create a new variable that is the career win/loss percentage (`Total_W/G_managed`). (Note: I did this after `collect()` otherwise the column type has to be set. . . )

• Sort the resulting data by the win/loss percentage variable (from largest to smallest).

```{r}
# will create an object first
manager_win_pct <- managers |>
  collect() |>
  mutate(
    Win_Pct = Total_W / G_managed
  ) |>
  arrange(desc(Win_Pct))

manager_win_pct
```

#### 17. Use the results from the previous two questions to answer this one! Return information for only the people that managed a team and were inducted into the hall of fame (regardless of their `category` in the hall of fame - you want to return all people in the hall of fame that ever managed at all). (Just return all of the variables from both of the above two questions - 9 total columns)

```{r}
hall_of_fame_managers <- hall_of_fame_names |>
  inner_join(manager_win_pct, by = "playerID")

hall_of_fame_managers
```

