---
title: "Homework 4"
authors: "Alfonso Straffon, Daria Smyslova, Michael Costello & Maya Szymborski"
format: html
editor: source
---

## Task 1: Conceptual Questions

**1. Rewrite the following nested function call using baseR’s chaining operator:**

```{r}
library(tidyverse)
arrange(filter(select(as_tibble(iris), starts_with("Petal"), Species), Petal.Length <
    1.55), Species)

# Recreate the nested function using baseR's chaining operator

iris |>
  as_tibble() |>
  select(starts_with("Petal"), Species) |>
  filter(Petal.Length < 1.55) |>
  arrange(Species)
```

**2. What is meant by long format data and wide format data? Which do we generally prefer for statistical analysis?**

*Long format* data has one observation per row, with different variables or measurements stored in separate columns. *Wide format* data has multiple measurements for the same observation stored across different columns, so the same observation may have values in several columns. Long format data is preferred for statistical analysis.

## Task 2:

*Note: Use chaining where possible!*

### Glass Data

In HW 3 you read in the glass data from https://www4.stat.ncsu.edu/\~online/datasets/glass.data. Let’s reread that data (use your code from HW 3) that also names the columns. Recall the last variable being `Type of Glass` with values of:– 1 building_windows_float_processed, – 2 building_windows_non_float_processed, – 3 vehicle_windows_float_processed, – 4 vehicle_windows_non_float_processed (none in this database), – 5 containers, – 6 tableware, – 7 headlamps.

```{r}
# use `read_csv()` from the `readr` package to import data
# use `col_names` to name the columns

glass_data <- readr::read_csv("https://www4.stat.ncsu.edu/online/datasets/glass.data", 
                       col_names = c("Id", "RI", "Na", "Mg", "Al", "Si", "K", "Ca", 
                                     "Ba", "Fe", "Type_of_Glass"))

# print tibble
glass_data
```

**3. Start a chain on your data object that overwrites the `Type_of_glass` variable using `mutate()`.**

• Create a `factor` version of the variable that replaces the numbers with string labels (that is, replace 1 with “building_windows_float_processed”, 2 with “building_win. . . ”, etc. - see the variable descriptions above to give meaningful values).

```{r}
# overwrite "Type_of_Glass" variable using `mutate()`
# replace numbers with string labels using `factor()`

glass_data <- glass_data |>
  mutate(
    Type_of_Glass = factor(
    Type_of_Glass,
    levels = c(1,2,3,4,5,6,7),
    labels = c(
      "building_windows_float_processed",
      "building_windows_non_float_processed",
      "vehicle_windows_float_processed",
      "vehicle_windows_non_float_processed",
      "containers",
      "tableware",
      "headlamps"
      )
    )
  )

# check 
levels(glass_data$Type_of_Glass)

```

**4. Continue your chain and keep only observations where the `Fe` variable is less than 0.2 and the `Type of Glass` is either “tableware” or “headlamp”.**

```{r}
# use `filter()` to only keep observations 
# where `Fe` < 0.2 and `Type_of_Glass` is either "tableware" or "headlamp" 

glass_data <- glass_data |>
  mutate(
    Type_of_Glass = factor(
    Type_of_Glass,
    levels = c(1,2,3,4,5,6,7),
    labels = c(
      "building_windows_float_processed",
      "building_windows_non_float_processed",
      "vehicle_windows_float_processed",
      "vehicle_windows_non_float_processed",
      "containers",
      "tableware",
      "headlamps"
      )
    )
  ) |>
  filter(
    Fe < 0.2,
    Type_of_Glass == "tableware" |
      Type_of_Glass == "headlamps"
  )

# check
glass_data
```

### Yeast Data

In HW 3 you read in the Yeast data from https://www4.stat.ncsu.edu/\~online/datasets/yeast.data. Read this data into `R` directly from the URL, adding column names as well.

```{r}
# use `read_table` from `readr` package to import data

yeast_data <- readr::read_table("https://www4.stat.ncsu.edu/~online/datasets/yeast.data", 
                         col_names = c("seq_name", "mcg", "gvh", "alm", "mit", 
                                       "erl", "pox", "vac", "nuc", "class"))

# check tibble
yeast_data
```

**5. Start a chain that removes the `seq_name` and `nuc` columns.**

**6. Continue your chain to add columns corresponding to the mean and median of each numeric variable (`mcg`, `gvh`, `alm`, `mit`, `erl`, `pox`, and `vac`) at each `class` grouping (see the `across()` function as we did in the `dplyr` video!).**

```{r}
# Questions 5-6

# remove `seq_name` and `nuc` columns using `select(-x)` 
# where x is the variable you want to remove

# then add mean and median columns corresponding to each numeric variable at 
# each `class` grouping using `group_by()` and `across()`

yeast_data <- yeast_data |>
  select(-seq_name, -nuc) |>
  group_by(class) |>
  mutate(
    across(
      mcg:vac,
      list(mean = mean, median = median)
      )
  )

names(yeast_data)
```

### Wine Data

In HW 3 you read in two data sets on wine. Read those two in and combine them as was done there.

```{r}
# Rear in white wine data
white_wine_data <- readxl::read_excel("data/white-wine.xlsx", sheet = 1)
white_wine_data_2 <- readxl::read_excel("data/white-wine.xlsx", sheet = 2)

# replace column names
colnames(white_wine_data) <- white_wine_data_2[[1]]

# add wine type
white_wine_data$type <- "white"

# import red wine data
red_wine_data <- readr::read_delim(
  "https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv", 
  delim = ";")

# replace column names
colnames(red_wine_data) <- white_wine_data_2[[1]] 

# add wine type
red_wine_data$type <- "red" 

# combine the two data sets 
wine_data <- dplyr::bind_rows(white_wine_data, red_wine_data) 

# check
table(wine_data$type)
```

**7. Start a chain on your new combined data object to filter the data to only look at high-quality wines (`quality` \> 6.5) and wines that have a reasonable alcohol value (`alcohol` \< 13.2).**

**8. Continue your chain to now sort the data from highest quality to lowest.**

**9. Continue your chain to select only the variables that contain `acid`, the `alcohol` variable, the `type` variable, and the `quality` variable.**

**10. Continue your chain to add the mean and standard deviation of the `alcohol` variable to the data set for each setting of the `quality` variable.**

```{r}
# Questions 7-10

# use `filter()`to only show wines with quality > 6.5 and alcohol < 13.2 

# then use `arrange()` to sort data from highest quality to lowest 

# then select only variables containing "acid" in name, 
# as well as the `alcohol`, `type`, and  `quality` variables 

# then add mean and standard deviation columns for the `alcohol` variable

# then group by `quality` variable 

wine_data <- wine_data |>
  filter(quality > 6.5, alcohol < 13.2) |>
  arrange(desc(quality)) |>
   select(
    contains("acid"),
    alcohol, 
    type,
    quality
  ) |>
  group_by(quality) |>
  mutate(
    mean_alcohol = mean(alcohol),
    sd_alcohol = sd(alcohol)
  )

# check
wine_data
```

## Task 3: Database Practice

Download the `Lahman.db` file associated with the *Connecting to Databases* notes/video. 

**11. Connect to the database and then look at all of the tables in the database.**

```{r}
# install.packages("RSQLite") in console
library(RSQLite)
library(DBI)

# connect to database & store connection
con <- dbConnect(
  RSQLite::SQLite(), 
  "data/lahman.db")

# look at all tables in database
dbListTables(con)
```

**12. Use the `tbl()` function and `dplyr` to *return* all of the data from the `Teams` table for the year 2015.**

```{r}
# install.packages("dplyr") in console
library(dplyr)

# use `tbl()` to return data from the `Teams` table for the year 2015

tbl(con, "Teams") |>
  filter(yearID == 2015)
```

**13. Repeat the above by using the `sql()` function within `tbl()` (here you have to write actual SQL!).**

```{r}
# use `sql()` to return data from the `Teams` table for the year 2015

tbl(
  con,
  sql("SELECT * 
      FROM Teams 
      WHERE yearID = 2015")
  )
```

**14. Return all of the players in the hall of fame, the year they were voted into the hall of fame, and their category (only those three variables!). See the `HallOfFame` table, the inducted variable is important here.**

```{r}
# return all players from the `HallOfFame` table, 
# plus the year they were voted in and their category

# dplyr
hall_of_fame <- tbl(con, "HallOfFame") |>
  filter(inducted == "Y") |>
  select(playerID, yearID, category)

# check data frame
hall_of_fame
```

**15. Combine the table from the previous question with the `People` table in order to have the `nameFirst` and `nameLast` names added to what was returned above.**

```{r}
# return first and last name of players from the `People` table

players <- tbl(con, "People") |>
  select(playerID, nameFirst, nameLast)

# check data frame
players

# combine with `hall_of_fame` table using `left_join()`

HOF_names <- players |>
  left_join(players, by = "playerID") |>
  collect()

# check tibble
HOF_names
```

**16. Return only the `playerID` (manager ID in this case), `G`, `W`, and `L` columns from the `Managers` table.**

• Use chaining to determine the overall win/loss records (sum of wins and sum of losses) for each of these hall of fame managers. – We haven’t quite covered this so I’ll help out. We want to use `group_by(`) with `summarize()` rather than `mutate()`. Something like this:

```{r}
# using `group_by()` and `summarize()` to
# return `playerID`, `G`, `W`, and `L` from the `Managers` table
# and determine overall win/loss records for each manager 

managers <- tbl(con, "Managers") |>
  select(playerID, G, W, L) |>
  group_by(playerID) |>
  summarize(G_managed = sum(G, na.rm = TRUE),
            Total_W = sum(W, na.rm = TRUE),
            Total_L = sum(L, na.rm = TRUE)
  )

# check
managers
```

• Create a new variable that is the career win/loss percentage (`Total_W/G_managed`). (Note: I did this after `collect()` otherwise the column type has to be set. . . )

• Sort the resulting data by the win/loss percentage variable (from largest to smallest).

```{r}
# create new variable that is career win/loss percentage (WL_PCT) using mutate()

# then sort data by the win/loss percentage (WL_PCT) variable 
# from largest to smallest using arrange()

manager_records <- managers |>
  collect() |>
  mutate(WL_PCT = Total_W / G_managed) |>
  arrange(desc(WL_PCT))

# check tibble
manager_records
```

**17. Use the results from the previous two questions to answer this one! Return information for only the people that managed a team and were inducted into the hall of fame (regardless of their `category` in the hall of fame - you want to return all people in the hall of fame that ever managed at all). (Just return all of the variables from both of the above two questions - 9 total columns)**

```{r}
# use `inner_join()` to return information for people that managed a team 
# AND were inducted into the Hall of Fame

HOF_managers <- HOF_names |>
  inner_join(manager_records, by = "playerID") 

# check
HOF_managers
```
