---
title: "Homework 4"
author: "Daria Smyslova"
format: html
editor: source
---

### Task 1: Conceptual Questions

#### 1. Rewrite the following nested function call using baseR’s chaining operator:

```{r}
library(tidyverse)

iris |>
  as_tibble() |>
  select(starts_with("Petal"), Species) |>
  filter(Petal.Length < 1.55) |>
  arrange(Species)
```

#### 2. What is meant by long format data and wide format data? Which do we generally prefer for statistical analysis?

**Wide format** data has one row for each observation and different measurements are stored in separate columns. **Long format** data has multiple rows for each observation, with one column identifying the variable or condition and another column containing its value. For statistical analysis, we generally prefer long format because it is easier to work with for modeling, grouping, and visualization.

### Task 2:

```{r}
library(tidyverse)

glass <- read_csv(
  "https://www4.stat.ncsu.edu/~online/datasets/glass.data",
  col_names = FALSE
)

names(glass) <- c(
  "Id", "RI", "Na", "Mg", "Al", "Si",
  "K", "Ca", "Ba", "Fe", "Type"
)

glass
```

3.  Start a chain on your data object that overwrites the Type_of_glass variable using `mutate()`.

    Create a `factor` version of the variable that replaces the numbers with string labels (that is, replace 1 with “building_windows_float_processed”, 2 with “building_win. . . ”, etc. - see the variable descriptions above to give meaningful values).

```{r}
glass <- glass |>
  mutate(
    Type = factor(
      Type,
      levels = c(1, 2, 3, 4, 5, 6, 7),
      labels = c(
        "building_windows_float_processed",
        "building_windows_non_float_processed",
        "vehicle_windows_float_processed",
        "vehicle_windows_non_float_processed",
        "containers",
        "tableware",
        "headlamps"
      )
    )
  )

glass
```

4\. Continue your chain and keep only observations where the `Fe` variable is less than 0.2 and the `Type of Glass` is either “tableware” or “headlamp”.

```{r}
glass <- glass |>
  filter(
    Fe < 0.2,
    Type %in% c("tableware", "headlamps")
  )

glass
```

```{r}
library(tidyverse)
glass_data <- read_csv("https://www4.stat.ncsu.edu/~online/datasets/glass.data",   col_names = 
  c("Id", "RI", "Na", "Mg", "Al", "Si", "K", "Ca", "Ba", "Fe", "TypeOfGlass")
)
```

#### 5. **Yeast data** In HW 3 you read in the Yeast data from https://www4.stat.ncsu.edu/\~online/datasets/yeast.data. Read this data into `R` directly from the URL, adding column names as well. Start a chain that removes the `seq_name` and `nuc` columns.

```{r}
yeast <- read_table(
  "https://www4.stat.ncsu.edu/~online/datasets/yeast.data",
  col_names = c(
    "seq_name", "mcg", "gvh", "alm", "mit",
    "erl", "pox", "vac", "nuc", "class"
  )
)

yeast
```

#### 6. Continue your chain to add columns corresponding to the mean and median of each numeric variable (`mcg`, `gvh`, `alm`, `mit`, `erl`, `pox`, and `vac`) at each `class` grouping (see the `across()` function as we did in the `dplyr` video!).

```{r}
yeast <- yeast |>
  select(-seq_name, -nuc) |>
  group_by(class) |>
  mutate(
    across(
      c(mcg, gvh, alm, mit, erl, pox, vac),
      list(mean = mean, median = median),
      .names = "{.col}_{.fn}"
    )
  )

yeast
```

#### 7. **Wine Data** In HW 3 you read in two data sets on wine. Read those two in and combine them as was done there. 

```{r}
#Read in White wine data
library(readxl)
wine_data <- read_excel("data/white-wine.xlsx")
#Add in type = White
wine_data <- wine_data |> mutate(type = "white")

#Read in the Red wine data
wine_data2 <- 
  read_delim("https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv", 
  delim = ";"
)
#Add in type = Red
wine_data2 <- wine_data2 |> mutate(type = "red")

#Combine the two dataset together
wine_data_final <- dplyr::bind_rows(wine_data, wine_data2)

#Read in and apply the Wine Labels (Which is missing "type", so we have to add that)
wine_data_lab <- read_excel("data/white-wine.xlsx", sheet =2)
wine_data_lab <- wine_data_lab |> add_row(Variables = "type")
colnames(wine_data_final) <- wine_data_lab[[1]]
```

#### Start a chain on your new combined data object to filter the data to only look at high-quality wines (`quality` \> 6.5) and wines that have a reasonable alcohol value (`alcohol` \< 13.2).

#### 8. Continue your chain to now sort the data from highest quality to lowest.

#### 9. Continue your chain to select only the variables that contain `acid`, the `alcohol` variable, the `type` variable, and the `quality` variable.

#### 10. Continue your chain to add the mean and standard deviation of the `alcohol` variable to the data set for each setting of the `quality` variable.

```{r}
wine_data_final <- wine_data_final |>
  
  # 7.Keep high-quality wines with alcohol less than 13.2
  filter(quality > 6.5, alcohol < 13.2) |>
  
  # 8.Sort from highest to lowest quality
  arrange(desc(quality)) |>
  
  # 9.Keep acid variables, alcohol, type, and quality
  select(contains("acid"), alcohol, type, quality) |>
  
  # 10.Add mean and standard deviation of alcohol for each quality
  group_by(quality) |>
  mutate(
    alcohol_mean = mean(alcohol),
    alcohol_sd = sd(alcohol)
  )

wine_data_final
```

### Task 3: Database Practice

#### 11. Download the `Lahman.db` file associated with the *Connecting to Databases* notes/video. Connect to the database and then look at all of the tables in the database.

```{r}
library(DBI)
library(RSQLite)
library(dplyr)

con <- dbConnect(
  RSQLite::SQLite(),
  "data/Lahman.db"
)

dbListTables(con)
```

#### 12. Use the `tbl()` function and `dplyr` to **return** all of the data from the `Teams` table for the year 2015.

```{r}
tbl(con, "Teams") |>
  filter(yearID == 2015)
```

#### 13. Repeat the above by using the `sql()` function within `tbl()` (here you have to write actual SQL!).

```{r}
tbl(
  con,
  sql("SELECT * FROM Teams WHERE yearID = 2015")
)
```

-   Note: Either use dplyr or write SQL queries for the rest of these questions!\*

#### 14. Return all of the players in the hall of fame, the year they were voted into the hall of fame, and their category (only those three variables!). See the `HallOfFame` table, the inducted variable is important here.

```{r}
hall_fame <- tbl(con, "HallOfFame") |>
  filter(inducted == "Y") |>
  select(playerID, yearID, category)

hall_fame
```

#### 15. Combine the table from the previous question with the `People` table in order to have the `nameFirst` and `nameLast` names added to what was returned above.

```{r}
hall_fame_names <- hall_fame |>
  left_join(
    tbl(con, "People") |>
      select(playerID, nameFirst, nameLast),
    by = "playerID"
  )

hall_fame_names
```

#### 16. Return only the `playerID` (manager ID in this case), `G`, `W`, and `L` columns from the `Managers` table. Use chaining to then:

• Determine the overall win/loss records (sum of wins and sum of losses) for each of these hall of fame managers. – We haven’t quite covered this so I’ll help out. We want to use `group_by(`) with `summarize()` rather than `mutate()`. Something like this:

```{r}
tbl(con, "Managers") |>
select(...) |>
group_by(playerID) |>
summarize(G_managed = sum(G, na.rm = TRUE),
		Total_W = sum(W, na.rm = TRUE),
		Total_L = sum(L, na.rm = TRUE))
```

• Create a new variable that is the career win/loss percentage (`Total_W/G_managed`). (Note: I did this after `collect()` otherwise the column type has to be set. . . )

• Sort the resulting data by the win/loss percentage variable (from largest to smallest).

```{r}
manager_records <- tbl(con, "Managers") |>
  select(playerID, G, W, L) |>
  group_by(playerID) |>
  summarize(
    G_managed = sum(G, na.rm = TRUE),
    Total_W = sum(W, na.rm = TRUE),
    Total_L = sum(L, na.rm = TRUE)
  )

manager_records
```

#### 17. Use the results from the previous two questions to answer this one! Return information for only the people that managed a team and were inducted into the hall of fame (regardless of their `category` in the hall of fame - you want to return all people in the hall of fame that ever managed at all). (Just return all of the variables from both of the above two questions - 9 total columns)

```{r}
hall_fame_managers <- hall_fame_names |>
  inner_join(
    manager_records,
    by = "playerID"
  )

hall_fame_managers
```
